---
title: "Homework2_2"
author: "Valeriia"
date: "31 03 2020"
output:
  html_document:
    df_print: paged
---

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(ggplot2)
library(reshape2)
library(data.table)
library(broom)
library(olsrr)
library(corrplot)
library(plotly)
```

Import data

```{r}
setwd("D:/Project")
air <- read.csv("Air.csv", sep=";", dec=",")
air <- air[,-c(16, 17)]
```

Make it long and check data.
We can see outliers in -200 - replace it with NA. We can see that one of columns - Non Metanic HydroCarbons concentration have a lot of NAs. Let's drop it and then drop NAs. We saved data loosing one column.
Also we can notice that not all columns have normal distribution

```{r message=FALSE}
str(air)
for (i in 3:ncol(air)) {
  air[,i][which(air[,i] == -200)] <- NA
}
air_long <- melt(air)
air_long <- air_long[,-c(1,2)]
ggplot(air_long, aes(value)) + 
  geom_histogram() + 
  facet_wrap(~variable, scales = "free")
air <- air[,c(1:4,6:15)]
air <- drop_na(air)
air_long <- melt(air)
air_long <- air_long[,-c(1,2)]
summary(air)
```

So, we need to normolize our data.
```{r}
normalize <- function(x) {
return ((x - min(x)) / (max(x) - min(x)))
}
air_norm <- as.data.frame(lapply(air[3:14], normalize))
air_long <- melt(air_norm)
```

Check our data. Quantile - the proportion of cases that are less than certain values. If the requirements of "normality" are here - the data should lie on a diagonal line.
We can see that several columns do not have normal distribution.

```{r}
#CO.GT
qqnorm(air_norm$CO.GT., pch = 1, frame = FALSE)
qqline(air_norm$CO.GT., col = "steelblue", lwd = 2)
#PT08.S1.CO
qqnorm(air_norm$PT08.S1.CO., pch = 1, frame = FALSE)
qqline(air_norm$PT08.S1.CO., col = "steelblue", lwd = 2)
#C6H6.GT
qqnorm(air_norm$C6H6.GT., pch = 1, frame = FALSE)
qqline(air_norm$C6H6.GT., col = "steelblue", lwd = 2)
#PT08.S2.NMHC
qqnorm(air_norm$PT08.S2.NMHC., pch = 1, frame = FALSE)
qqline(air_norm$PT08.S2.NMHC., col = "steelblue", lwd = 2)
#NOx.GT
qqnorm(air_norm$NOx.GT., pch = 1, frame = FALSE)
qqline(air_norm$NOx.GT., col = "steelblue", lwd = 2)
#PT08.S3.NOx
qqnorm(air_norm$PT08.S3.NOx., pch = 1, frame = FALSE)
qqline(air_norm$PT08.S3.NOx., col = "steelblue", lwd = 2)
#NO2.GT
qqnorm(air_norm$NO2.GT., pch = 1, frame = FALSE)
qqline(air_norm$NO2.GT., col = "steelblue", lwd = 2)
#PT08.S4.NO2
qqnorm(air_norm$PT08.S4.NO2., pch = 1, frame = FALSE)
qqline(air_norm$PT08.S4.NO2., col = "steelblue", lwd = 2)
#PT08.S5.O3
qqnorm(air_norm$PT08.S5.O3., pch = 1, frame = FALSE)
qqline(air_norm$PT08.S5.O3., col = "steelblue", lwd = 2)
#T
qqnorm(air_norm$T, pch = 1, frame = FALSE)
qqline(air_norm$T, col = "steelblue", lwd = 2)
#RH
qqnorm(air_norm$RH, pch = 1, frame = FALSE)
qqline(air_norm$RH, col = "steelblue", lwd = 2)
#AH
qqnorm(air_norm$AH, pch = 1, frame = FALSE)
qqline(air_norm$AH, col = "steelblue", lwd = 2)
```

Response C6H6.GT. We can see columns that have linear response to C6H6.GT.

```{r}
mod <- lm(data = air_norm, C6H6.GT. ~ CO.GT.)
summary(mod)
augment(mod)%>%
   ggplot(aes(x = CO.GT., y = C6H6.GT.))+
  geom_point()+
  geom_line(aes(y= .fitted), color = "blue", size = 1)
residuals(mod) %>% hist(main = "Residuals CO.GT.")
plot(mod, which = 1)
mod1 <- lm(data = air_norm, C6H6.GT. ~ PT08.S1.CO.)
summary(mod1)
augment(mod1)%>%
   ggplot(aes(x = PT08.S1.CO., y = C6H6.GT.))+
  geom_point()+
  geom_line(aes(y= .fitted), color = "blue", size = 1)
residuals(mod1) %>% hist(main = "Residuals PT08.S1.CO.")
plot(mod1, which = 1)
mod2 <- lm(data = air_norm, C6H6.GT. ~ PT08.S2.NMHC.)
summary(mod2)
augment(mod2)%>%
   ggplot(aes(x = PT08.S2.NMHC., y = C6H6.GT.))+
  geom_point()+
  geom_line(aes(y= .fitted), color = "blue", size = 1)
residuals(mod2) %>% hist(main = "Residuals PT08.S2.NMHC.")
plot(mod2, which = 1)
mod3 <- lm(data = air_norm, C6H6.GT. ~ NOx.GT.)
summary(mod3)
augment(mod3)%>%
   ggplot(aes(x = NOx.GT., y = C6H6.GT.))+
  geom_point()+
  geom_line(aes(y= .fitted), color = "blue", size = 1)
residuals(mod3) %>% hist(main = "Residuals NOx.GT.")
plot(mod3, which = 1)
mod4 <- lm(data = air_norm, C6H6.GT. ~ PT08.S3.NOx.)
summary(mod4)
augment(mod4)%>%
   ggplot(aes(x = PT08.S3.NOx., y = C6H6.GT.))+
  geom_point()+
  geom_line(aes(y= .fitted), color = "blue", size = 1)
residuals(mod4) %>% hist(main = "Residuals PT08.S3.NOx.")
plot(mod4, which = 1)
mod5 <- lm(data = air_norm, C6H6.GT. ~ NO2.GT.)
summary(mod5)
augment(mod5)%>%
   ggplot(aes(x = NO2.GT., y = C6H6.GT.))+
  geom_point()+
  geom_line(aes(y= .fitted), color = "blue", size = 1)
residuals(mod5) %>% hist(main = "Residuals NO2.GT.")
plot(mod5, which = 1)
mod6 <- lm(data = air_norm, C6H6.GT. ~ PT08.S4.NO2.)
summary(mod6)
augment(mod6)%>%
   ggplot(aes(x = PT08.S4.NO2., y = C6H6.GT.))+
  geom_point()+
  geom_line(aes(y= .fitted), color = "blue", size = 1)
residuals(mod6) %>% hist(main = "Residuals PT08.S4.NO2.")
plot(mod6, which = 1)
mod7 <- lm(data = air_norm,C6H6.GT.  ~ PT08.S5.O3.)
summary(mod7)
augment(mod7)%>%
   ggplot(aes(x = PT08.S5.O3., y = C6H6.GT.))+
  geom_point()+
  geom_line(aes(y= .fitted), color = "blue", size = 1)
residuals(mod7) %>% hist(main = "Residuals PT08.S5.O3.")
plot(mod7, which = 1)
mod8 <- lm(data = air_norm,C6H6.GT. ~ `T`)
summary(mod8)
augment(mod8)%>%
   ggplot(aes(x = `T`, y = C6H6.GT.))+
  geom_point()+
  geom_line(aes(y= .fitted), color = "blue", size = 1)
residuals(mod8) %>% hist(main = "Residuals T")
plot(mod8, which = 1)
mod9 <- lm(data = air_norm, C6H6.GT. ~ RH)
summary(mod9)
augment(mod9)%>%
   ggplot(aes(x = RH, y = C6H6.GT.))+
  geom_point()+
  geom_line(aes(y= .fitted), color = "blue", size = 1)
residuals(mod9) %>% hist(main = "Residuals RH")
plot(mod9, which = 1)
mod10 <- lm(data = air_norm, C6H6.GT. ~ AH)
summary(mod10)
augment(mod10)%>%
   ggplot(aes(x = AH, y = C6H6.GT.))+
  geom_point()+
  geom_line(aes(y= .fitted), color = "blue", size = 1)
residuals(mod10) %>% hist(main = "Residuals AH")
plot(mod10, which = 1)
```

```{r}
pairs(air_norm[,sapply(air_norm, is.double)])
```
Check correlation.

```{r}
corrplot.mixed(cor(air_norm, method = "kendall"), number.cex = .7)
```
Check multicolinearity. I decided to see what's going on by hands. For multic. we need to have CI>30 and VIF>10. Also, we can see it when coeficients of columns alone are not significant but together they give significant p-value.

```{r}
mod <-  lm(C6H6.GT.~PT08.S2.NMHC. + NOx.GT. + PT08.S3.NOx. + NO2.GT. + PT08.S4.NO2., data = air_norm)
summary(mod)
X <- model.matrix(~PT08.S2.NMHC. + NOx.GT. + PT08.S3.NOx. + NO2.GT. + PT08.S4.NO2., data = air_norm)
XX <- t(X) %*% X
eigen <- eigen(XX)
CI <- sqrt(max(eigen$values) / min(eigen$values))
CI
# Also can count it via formula 1/1-R^2
ols_coll_diag(mod)
residuals(mod) %>% hist(main = "Residuals PT08.S2.NMHC. + NOx.GT. + PT08.S3.NOx. + NO2.GT. + PT08.S4.NO2.")
plot(mod, which = 1)

mod <-  lm(C6H6.GT.~PT08.S2.NMHC. + NOx.GT. + PT08.S3.NOx. + NO2.GT., data = air_norm)
summary(mod)
X <- model.matrix(~~PT08.S2.NMHC. + NOx.GT. + PT08.S3.NOx. + NO2.GT., data = air_norm)
XX <- t(X) %*% X
eigen <- eigen(XX)
CI <- sqrt(max(eigen$values) / min(eigen$values))
CI
# Also can count it via formula 1/1-R^2
ols_coll_diag(mod)
residuals(mod) %>% hist(main = "Residuals PT08.S2.NMHC. + NOx.GT. + PT08.S3.NOx. + NO2.GT.")
plot(mod, which = 1)

mod <-  lm(C6H6.GT.~PT08.S2.NMHC. + NOx.GT. + PT08.S3.NOx., data = air_norm)
summary(mod)
X <- model.matrix(~PT08.S2.NMHC. + NOx.GT. + PT08.S3.NOx., data = air_norm)
XX <- t(X) %*% X
eigen <- eigen(XX)
CI <- sqrt(max(eigen$values) / min(eigen$values))
CI
# Also can count it via formula 1/1-R^2
ols_coll_diag(mod)
residuals(mod) %>% hist(main = "Residuals PT08.S2.NMHC. + NOx.GT. + PT08.S3.NOx.")
plot(mod, which = 1)

mod <-  lm(C6H6.GT.~PT08.S2.NMHC. + NOx.GT., data = air_norm)
summary(mod)
X <- model.matrix(~PT08.S2.NMHC. + NOx.GT., data = air_norm)
XX <- t(X) %*% X
eigen <- eigen(XX)
CI <- sqrt(max(eigen$values) / min(eigen$values))
CI
# Also can count it via formula 1/1-R^2
ols_coll_diag(mod)
residuals(mod) %>% hist(main = "Residuals PT08.S2.NMHC. + NOx.GT.")
plot(mod, which = 1)

mod <-  lm(C6H6.GT.~PT08.S2.NMHC. + PT08.S3.NOx. + NO2.GT. + PT08.S4.NO2., data = air_norm)
summary(mod)
X <- model.matrix(~PT08.S2.NMHC. + PT08.S3.NOx. + NO2.GT. + PT08.S4.NO2., data = air_norm)
XX <- t(X) %*% X
eigen <- eigen(XX)
CI <- sqrt(max(eigen$values) / min(eigen$values))
CI
# Also can count it via formula 1/1-R^2
ols_coll_diag(mod)
residuals(mod) %>% hist(main = "Residuals PT08.S2.NMHC. + PT08.S3.NOx. + NO2.GT. + PT08.S4.NO2.")
plot(mod, which = 1)

mod <-  lm(C6H6.GT.~ PT08.S3.NOx. + NO2.GT. + PT08.S4.NO2., data = air_norm)
summary(mod)
X <- model.matrix(~PT08.S3.NOx. + NO2.GT. + PT08.S4.NO2., data = air_norm)
XX <- t(X) %*% X
eigen <- eigen(XX)
CI <- sqrt(max(eigen$values) / min(eigen$values))
CI
# Also can count it via formula 1/1-R^2
ols_coll_diag(mod)
residuals(mod) %>% hist(main = "Residuals PT08.S3.NOx. + NO2.GT. + PT08.S4.NO2.")
plot(mod, which = 1)
```

Final model
Residuals should be normally distributed and the Q-Q Plot will show this. If residuals follow close to a straight line on this plot, it is a good indication they are normally distributed.
```{r}
# Good model but NOx.GT. not normally distributed
# I've decided not to transform data, because I don't want to see log-log transformation
mod <-  lm(C6H6.GT.~PT08.S2.NMHC. + NOx.GT. + PT08.S3.NOx. + NO2.GT., data = air_norm)
summary(mod)
X <- model.matrix(~PT08.S2.NMHC. + NOx.GT. + PT08.S3.NOx. + NO2.GT., data = air_norm)
XX <- t(X) %*% X
eigen <- eigen(XX)
CI <- sqrt(max(eigen$values) / min(eigen$values))
CI
ols_coll_diag(mod)
ols_plot_resid_fit_spread(mod)
ols_plot_obs_fit(mod)

mod <-  lm(C6H6.GT.~CO.GT. + PT08.S4.NO2., data = air_norm)
summary(mod)
X <- model.matrix(~CO.GT. + PT08.S4.NO2., data = air_norm)
XX <- t(X) %*% X
eigen <- eigen(XX)
CI <- sqrt(max(eigen$values) / min(eigen$values))
CI
ols_coll_diag(mod)
ols_plot_resid_fit_spread(mod)
ols_plot_obs_fit(mod)
```

Separated data in 75:25%
We can see that R^2 is close to 1 which says about linear dependence. P-value is close to 0 - independent variables explain the dynamics of the dependent variable.

```{r message=FALSE, warning=FALSE}
set.seed(2)
sep <- sample.int(n = nrow(air_norm), size = floor(.75*nrow(air_norm)))
train <- air_norm[sep,]
test <- air_norm[-sep,]
train1 <- train[,c('C6H6.GT.', 'CO.GT.', 'PT08.S4.NO2.')]
test1 <- test[,c('C6H6.GT.', 'CO.GT.', 'PT08.S4.NO2.')]
mod1 <- lm(C6H6.GT.~CO.GT. + PT08.S4.NO2., data = train1)
a1 <- summary(mod1)
pred1 <- predict(mod1, newdata = test1)
test1$C6H6.GT._pred <- pred1
head(test1)
test1 <- rbindlist(list(test1[,c(2,3,1)], test1[,c(2,3,4)]))
train1$type <- 'train'
test1$type <- 'test'
test1[1:(nrow(test1)/2),4] <- 'real value'
all <- rbind(train1, test1)
plot_ly(data = all,
        z = ~C6H6.GT.,
        y = ~PT08.S4.NO2.,
        x = ~CO.GT., opacity = 0.7, color = ~type)%>%
  layout(title = paste("R2", round(a1$r.squared, 3),
                                  sep = ": "),
                            paste("pvalue", 2.2e-16, sep = ": "),
         xaxis = list(title = "CO.GT.",
                      zeroline = FALSE),
         yaxis = list(title = "PT08.S4.NO2.",
                      zeroline = FALSE),
         zaxis = list(title = "C6H6.GT.",
                      zeroline = FALSE))
```