---
title: "HW2.1_Mary_Futey"
author: "Mary Futey"
date: "4/3/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(magrittr)
library(plyr)
library(reshape2)
library(corrplot)
library(data.table)
```

## Part 1: Anscombe’s dataset
*Scatter plot facetted by set

```{r Ans facet}
knitr::kable(anscombe)
anscombe.1 <- data.frame(x = anscombe[["x1"]], 
                         y = anscombe[["y1"]], 
                         Set = "Anscombe Set 1")
anscombe.2 <- data.frame(x = 
                           anscombe[["x2"]], 
                         y = anscombe[["y2"]], 
                         Set = "Anscombe Set 2")
anscombe.3 <- data.frame(x = anscombe[["x3"]], 
                         y = anscombe[["y3"]], 
                         Set = "Anscombe Set 3")
anscombe.4 <- data.frame(x = anscombe[["x4"]], 
                         y = anscombe[["y4"]], 
                         Set = "Anscombe Set 4")
anscombe.data <- rbind(anscombe.1, anscombe.2, anscombe.3, anscombe.4)

ans_facet <- ggplot(anscombe.data, 
                    aes(x = x, 
                        y = y))+ 
  geom_point(color = "black")+ 
  facet_wrap(~Set, ncol = 2)

ans_facet

```

*Summary calculation (mean, sd) grouped by set

```{r Ans mean and sd}

aggregate(cbind(x, y) ~ Set, anscombe.data, mean)
aggregate(cbind(x, y) ~ Set, anscombe.data, sd)

```

*Pearson’s correlation by set, and non-parametric, and p-value 

```{r Ans Pearsons, non-parametric, pval}

correlation <- function(data) {
  a <- data.frame(pearson = cor.test(data$x, data$y, method = "pearson")$statistic)
  b <- data.frame(p_pearson = cor.test(data$x, data$y, method = "pearson")$p.value) 
  c <- data.frame(kendall = cor.test(data$x, data$y, method = "kendall")$statistic)
  d <- data.frame(p_kendall = cor.test(data$x, data$y, method = "kendall")$p.value)
  e <- data.frame(spearman = cor.test(data$x, data$y, method = "spearman")$statistic)
  f <- data.frame(p_spearman = cor.test(data$x, data$y, method = "spearman")$p.value) 
    
  return(list(a, b, c, d, e, f))
}

res <- correlation(anscombe.data)
ldply(res)



```
*Add geom_smooth()totheplot

```{r Ans plot with geom_smooth}
ans_smooth <- ans_facet + 
  geom_smooth(formula = y ~ x, method = "lm", se = FALSE, data = anscombe.data)
ans_smooth

```

## Air Quality dataset

* Explore data set, clean if needed: convert date and time to factor, check for NAs, switch comma to decimal
* Explore each variable independently: remove "-200" values as they are not possible / erroneous 

```{r Air Q load, explore and clean}
airq <- read.csv("/Users/maryfutey/desktop/AirQualityUCI/AirQualityUCI.csv", 
                 header = TRUE,
                 dec=",")
head(airq)

airq_long <- gather(airq, key="measurement", value="value", -c(Date,Time))

airq_long$Date <- as.factor(airq_long$Date)
airq_long$Time <- as.factor(airq_long$Time)
airq_long$measurement <- as.factor(airq_long$measurement)

airq_fil <- airq_long %>% 
  filter_all(all_vars(. != -200))

head(airq_fil)
colSums(is.na(airq_fil))

p1 <- airq_fil %>% ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~measurement, scales = "free")

p1


```


* Need to normalize data

```{r Air Q norm}
airq_wide <- spread(airq_fil, key = "measurement" ,value = "value")
airq_wide <- na.omit(airq_wide)

norm <- function(x) {
    (x - min(x)) / (max(x) - min(x))
  }

airq_norm <- as.data.frame(lapply(airq_wide[3:15], norm))
airq_norm$Date <- airq_wide$Date
airq_norm$Time <- airq_wide$Time


airq_norm_long <- gather(airq_norm, key = "measurement", value = "value", -c(Date, Time))

p2 <- airq_norm_long %>%  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~measurement, scales = "free")
 
p2

```


* Cross correlations

```{r Air Q cross coor}
airq_sub <- airq_norm[, 1:13]
colnames(airq_sub) <- c(
  "Abs H",
  "C6H6",
  "CO",
  "NMHC",
  "NO2",
  "NOx",
  "S1.CO",
  "S2.NMHC",
  "S3.NOx",
  "S4.NO2",
  "S5.O3",
  "Rel H", 
  "Temp"
)

airq_cor <- round(cor(airq_sub, method = "kendall"),2)

#reorder, create upper triangle
reorder_airq_cor <- function(airq_cor){
  # Use correlation between variables as distance
  dd <- as.dist((1-airq_cor)/2)
  hc <- hclust(dd)
  airq_cor <-airq_cor[hc$order, hc$order]
}

get_upper_tri <- function(airq_cor){
  airq_cor[lower.tri(airq_cor)]<- NA
  return(airq_cor)
}

airq_cor <- reorder_airq_cor(airq_cor)
upper_tri <- get_upper_tri(airq_cor)
melt_uppertri <- melt(upper_tri, na.rm = TRUE)

p3 <- ggplot(melt_uppertri, 
             aes(Var2, Var1, 
                 fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Kendall\nCorrelation") +
  ggtitle("Correlation Matrix for Air Quality Dataset") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 7, hjust = 1)) +
  coord_fixed() +
  geom_text(aes(Var2, 
                Var1, 
                label = value), 
            color = "black", size = 2.8) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(0.6, 0.7),
    legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5))

p3



p4 <- corrplot.mixed(airq_cor, 
                     tl.col = "black", tl.cex = 0.6,
                     lower.col = "black", number.cex = 0.7)

p4
```


* Build simple linear models with each predictor, check assumptions, response C6H6


#Absolute humidity - data not linear
```{r Air Q AH}
ggplot(airq_norm, aes(x = AH,
                   y = C6H6.GT.)) +
  geom_point() +
  ggtitle("C6H6.GT. vs absolute humidity")
AH <- lm(C6H6.GT. ~ AH, airq_norm)
residuals(AH) %>% hist(main = "residuals absolute humidity")
summary(AH)

```

#Ave CO  (CO.GT.)

```{r Air Q CO}
ggplot(airq_norm, aes(x = CO.GT.,
                   y = C6H6.GT.)) +
  geom_point() +
  ggtitle("C6H6.GT. vs Ave CO")

CO <- lm(C6H6.GT. ~ CO.GT., airq_norm)
residuals(CO) %>% hist(main = "residuals ave CO")
summary(CO)

```

#Non Metanic HydroCarbons (NMHC.GT.)

```{r Air Q NMHC}
ggplot(airq_norm, aes(x = NMHC.GT.,
                   y = C6H6.GT.)) +
  geom_point() +
  ggtitle("C6H6.GT. vs Non Metanic HydroCarbons")

NMHC <- lm(C6H6.GT. ~ NMHC.GT., airq_norm)
residuals(NMHC) %>% hist(main = "residuals Non Metanic HydroCarbons")
summary(NMHC)
```

#ave NO2 (NO2.GT.)
```{r Air Q NO2}
ggplot(airq_norm, aes(x = NO2.GT.,
                   y = C6H6.GT.)) +
  geom_point() +
  ggtitle("C6H6.GT. vs averaged NO2")

NO2 <- lm(C6H6.GT. ~ NO2.GT., airq_norm)
residuals(NO2) %>% hist(main = "residuals ave NO2")
summary(NO2)
```

#ave NOx (NO2.GT.)
```{r Air Q NOx}
ggplot(airq_norm, aes(x = NOx.GT.,
                   y = C6H6.GT.)) +
  geom_point() +
  ggtitle("C6H6.GT. vs averaged NOx")

NOx <- lm(C6H6.GT. ~ NOx.GT., airq_norm)
residuals(NOx) %>% hist(main = "residuals ave NOx")
summary(NOx)
```

#ave tin oxide (PT08.S1.CO.)
```{r Air Q PT08.S1}
ggplot(airq_norm, aes(x = PT08.S1.CO.,
                   y = C6H6.GT.)) +
  geom_point() +
  ggtitle("C6H6.GT. vs tin oxide / PT08.S1.CO.")

PT08.S1 <- lm(C6H6.GT. ~ PT08.S1.CO., airq_norm)
residuals(PT08.S1) %>% hist(main = "residuals tin oxide")
summary(PT08.S1)
```

#ave titania (PT08.S2.NMHC)
```{r Air Q PT08.S2}
ggplot(airq_norm, aes(x = PT08.S2.NMHC.,
                   y = C6H6.GT.)) +
  geom_point() +
  ggtitle("C6H6.GT. vs titania / PT08.S2.NMHC")

PT08.S2 <- lm(C6H6.GT. ~ PT08.S2.NMHC., airq_norm)
residuals(PT08.S2) %>% hist(main = "residuals ave titania")
summary(PT08.S2)
```
#ave tungsten oxide NOx targeted (PT08.S3.NOx.) - data not linear
```{r Air Q PT08.S3}
ggplot(airq_norm, aes(x = PT08.S3.NOx.,
                   y = C6H6.GT.)) +
  geom_point() +
  ggtitle("C6H6.GT. vs tungsten oxide NOx / PT08.S3.NOx.")

PT08.S3 <- lm(C6H6.GT. ~ PT08.S3.NOx., airq_norm)
residuals(PT08.S3) %>% hist(main = " residuals tungsten oxide NOx")
summary(PT08.S3)
```
#ave tungsten oxide NO2 targeted (PT08.S4.NO2.) 
```{r Air Q PT08.S4}
ggplot(airq_norm, aes(x = PT08.S4.NO2.,
                   y = C6H6.GT.)) +
  geom_point() +
  ggtitle("C6H6.GT. vs tungsten oxide NO2 / PT08.S4.NO2")

PT08.S4 <- lm(C6H6.GT. ~ PT08.S4.NO2., airq_norm)
residuals(PT08.S4) %>% hist(main = " residuals tungsten oxide NO2")
summary(PT08.S4)
```
#ave indium oxide (PT08.S5.O3.)
```{r Air Q PT08.S5}
ggplot(airq_norm, aes(x = PT08.S5.O3.,
                   y = C6H6.GT.)) +
  geom_point() +
  ggtitle("C6H6.GT. vs indium oxide / PT08.S5.O3")

PT08.S5 <- lm(C6H6.GT. ~ PT08.S5.O3., airq_norm)
residuals(PT08.S5) %>% hist(main = "residuals ave indium oxide")
summary(PT08.S5)
```

#Relative humidity (RH) - data not linear
```{r Air Q RH}
ggplot(airq_norm, aes(x = RH,
                   y = C6H6.GT.)) +
  geom_point() +
  ggtitle("C6H6.GT. vs Relative humidity")

RH <- lm(C6H6.GT. ~ RH, airq_norm)
residuals(RH) %>% hist(main = "residuals relative humidity")
summary(RH)
```

#Temperature
```{r Air Q T}
ggplot(airq_norm, aes(x = T,
                   y = T)) +
  geom_point() +
  ggtitle("C6H6.GT. vs Temperature")

TP <- lm(C6H6.GT. ~ T, airq_norm)
residuals(TP) %>% hist(main = "residuals temperature")
summary(TP)

```


* For 2-3 of the models create train-test sets, plot the model, for the test set color real and predicted points differently; R^2 and p-value to title

```{r Air Q train test}

#ave CO and NO2
set.seed(88)
data <- airq_norm
sample <- sample.int(n = nrow(data), size = floor(.75*nrow(data)))
train <- data[sample, ]
test <- data[-sample, ]

train_CO <- train[,c("C6H6.GT.", "CO.GT." )]
test_CO <- test[,c("C6H6.GT.", "CO.GT." )]
new_mod_CO <- lm(data = train_CO, 
              C6H6.GT. ~ CO.GT.)


#summary
sum_CO <- summary(new_mod_CO)
print(sum_CO)

#use model, predicts new model from newdata (test)
#predictions for new test data set, add new column
pred_CO <- predict(new_mod_CO, newdata = test_CO)
test$CO.GT._pred <- pred_CO 


#make combined dataset
train_CO$data <- "train"
test_CO$data <- "test"
test_CO[1:(nrow(test_CO)/2),3] <- "real"
comb_CO <- rbind(train_CO, test_CO)

#trained and test should be similar shape
ggplot(data = comb_CO,
       aes(x = CO.GT.,
           y = C6H6.GT.,
           color = data)) +
  geom_point() +
  geom_smooth(method = "lm", 
              color = "black") +
  ggtitle(paste("R2", round(sum_CO$r.squared, 3),
                sep = ": "),
          paste("pvalue <2e-16" ))

#same for NO2
train_NO2 <- train[,c("C6H6.GT.", "NO2.GT." )]
test_NO2 <- test[,c("C6H6.GT.", "NO2.GT." )]
mod_NO2 <- lm(data = train_NO2, 
              C6H6.GT. ~ NO2.GT.)

sum_NO2 <- summary(mod_NO2)
print(sum_NO2)

pred_NO2 <- predict(mod_NO2, newdata = test_NO2)
test$NO2.GT._pred <- pred_NO2

train_NO2$data <- "train"
test_NO2$data <- "test"
test_NO2[1:(nrow(test_NO2)/2),3] <- "real"
comb_NO2 <- rbind(train_NO2, test_NO2)

ggplot(data = comb_NO2,
       aes(x = NO2.GT.,
           y = C6H6.GT.,
           color = data)) +
  geom_point() +
  geom_smooth(method = "lm", 
              color = "black") +
  ggtitle(paste("R2", round(sum_CO$r.squared, 3),
                sep = ": "),
          paste("pvalue <2e-16" ))
