---
title: "HW2.2_Mary_Futey"
author: "Mary Futey"
date: "4/26/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(magrittr)
library(plyr)
library(reshape2)
library(corrplot)
library(data.table)
library(car)

```

## Air Quality dataset

* Convert date and time to factor, check for NAs, switch comma to decimal

```{r Air Q load}
airq <- read.csv("/Users/maryfutey/desktop/AirQualityUCI/AirQualityUCI.csv", 
                 header = TRUE,
                 dec=",")
#deleted column  NMHC.GT. as there were many "-200" values
airq <- airq[, -5]

summary(airq)

airq_long <- gather(airq, key="measurement", value="value", -c(Date,Time))

airq_long$Date <- as.factor(airq_long$Date)
airq_long$Time <- as.factor(airq_long$Time)
airq_long$measurement <- as.factor(airq_long$measurement)

```

* Clean and remove "-200" values as they are not possible / erroneous 

```{r Air Q  clean}
airq_fil <- airq_long %>% 
  filter(value != -200)

summary(airq_fil)
colSums(is.na(airq_fil))

```

* Need to normalize data

```{r Air Q norm}
airq_wide <- spread(airq_fil, key = "measurement" ,value = "value")
airq_wide <- na.omit(airq_wide)

norm <- function(x) {
    (x - min(x)) / (max(x) - min(x))
  }

airq_norm <- as.data.frame(lapply(airq_wide[3:14], norm))
airq_norm$Date <- airq_wide$Date
airq_norm$Time <- airq_wide$Time

airq_norm_long <- gather(airq_norm, key = "measurement", value = "value", -c(Date, Time))

```

#Explore multicollinearity 

*Choose good predictors 

*Check residuals 

*Do you need some non-linear transformations for some of predictors? Or maybe response? 
+log(x)
+sqrt(x)
+(x)^2

** selected two predictor with high R^2, but non-linear tp transform: PT08.S2.NMHC.


```{r Air non-linear transformation PT08.S2.NMHC.}

# PT08.S2.NMHC.: residuals look better after quadratic transformtion, 
#however residuals vs fitted plot cubic
# R^2 of 0.999, how so high?
lm_3 <- lm(C6H6.GT. ~ PT08.S2.NMHC., airq_norm)
residuals(lm_3) %>% hist(main = "residuals PT08.S2.NMHC.")
summary(lm_3)
plot(lm_3, which = 1)

airq_norm$NMHC_sq <- (airq_norm$PT08.S2.NMHC.)^2

lm_4 <- lm(C6H6.GT. ~ PT08.S2.NMHC. + NMHC_sq, airq_norm)
residuals(lm_4) %>% hist(main = "residuals PT08.S2.NMHC.^2")
summary(lm_4)
plot(lm_4, which = 1)

airq_norm$NMHC_rt <- sqrt(airq_norm$PT08.S2.NMHC.)

lm_2 <- lm(C6H6.GT. ~ PT08.S2.NMHC. + NMHC_rt, airq_norm)
residuals(lm_2) %>% hist(main = "residuals PT08.S2.NMHC. sqrt transformed")
summary(lm_2)
plot(lm_2, which = 1)


```

## Make a coorelation matrix to select predictors to use in multiple linear regression
* due to deletion of NHMC column, matrix is differnet from the previous homework 

```{r coorelation matrix}

airq_sub <- airq_norm[, 1:12]
colnames(airq_sub) <- c(
  "Abs H",
  "C6H6",
  "CO",
  "NO2",
  "NOx",
  "S1.CO",
  "S2.NMHC",
  "S3.NOx",
  "S4.NO2",
  "S5.O3",
  "Rel H", 
  "Temp"
)

airq_cor <- round(cor(airq_sub, method = "kendall"),2)

#reorder, create upper triangle
reorder_airq_cor <- function(airq_cor){
  # Use correlation between variables as distance
  dd <- as.dist((1-airq_cor)/2)
  hc <- hclust(dd)
  airq_cor <-airq_cor[hc$order, hc$order]
}

get_upper_tri <- function(airq_cor){
  airq_cor[lower.tri(airq_cor)]<- NA
  return(airq_cor)
}

airq_cor <- reorder_airq_cor(airq_cor)
upper_tri <- get_upper_tri(airq_cor)
melt_uppertri <- melt(upper_tri, na.rm = TRUE)

p1 <- ggplot(melt_uppertri, 
             aes(Var2, Var1, 
                 fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Kendall\nCorrelation") +
  ggtitle("Correlation Matrix for Air Quality Dataset") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 7, hjust = 1)) +
  coord_fixed() +
  geom_text(aes(Var2, 
                Var1, 
                label = value), 
            color = "black", size = 2.6) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(0.6, 0.7),
    legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5))

p1
```

```{r Air CO predictor }

#single linear regression with good predictor for baseline comparison 

lm_5 <- lm(C6H6.GT. ~ CO.GT., airq_norm)
summary(lm_5)
residuals(lm_5) %>% hist(main = "residuals ave CO")

ggplot(lm_5, 
       aes(x = CO.GT., y = C6H6.GT.)) +
  geom_point() +
  geom_line(aes(y = .fitted), color = "red", size = 1) +
  ggtitle("predictor ave CO")

```

## selected the following predictors: CO.GT., PT08.S1.CO. and PT08.S2.NMHC. for building a model

```{r Air multiple regression 3 predictors}

#CO: corr 0.88, good residuals, high R^2, sigif, linear
#PT08.S1.CO., corr 0.83, residuals okay, high R^2, sigif, linear
#PT08.S2.NMHC., sym residuals, high R^2, sigif, linear 

#very high R^2, but residuals vs fitted plot not linear

lm_6 <- lm(C6H6.GT. ~ CO.GT. + PT08.S1.CO. + PT08.S2.NMHC., airq_norm)
summary(lm_6)
residuals(lm_6) %>% hist(main = "residuals multi regression 3 predictors and S2.NMHC transformed")

plot(lm_6, which = 1)


```


```{r Air multiple regression 3 predictors and transformed}

#CO: corr 0.88, good residuals, high R^2, sigif, linear
#PT08.S1.CO., corr 0.83, residuals okay, high R^2, sigif, linear
#PT08.S2.NMHC., sym residuals, high R^2, sigif, linear 

#inclusion of transformed PT08.S2.NMHC appears to make residuals vs fitted plot cubic?

lm_6 <- lm(C6H6.GT. ~ CO.GT. + PT08.S1.CO. + PT08.S2.NMHC. + NMHC_sq, airq_norm)
summary(lm_6)
residuals(lm_6) %>% hist(main = "residuals multi regression 3 predictors and S2.NMHC transformed")

plot(lm_6, which = 1)


```

## Model with the following predictors: PT08.S1.CO. and PT08.S2.NMHC.

```{r Air multiple regression 2 predictors PT08.S1.CO. and PT08.S2.NMHC.}

# residuals vs fitted plot is not linear
lm_7 <- lm(C6H6.GT. ~ PT08.S1.CO. + PT08.S2.NMHC., airq_norm)
summary(lm_7)
residuals(lm_7) %>% hist(main = "residuals multi PT08.S1.CO. + PT08.S2.NMHC. predictors")

plot(lm_7, which = 1)

# variance inflation factor: 
#how much the variance of a coefficient is inflated due to multicollinearity 
#vif <5, which is good, however based on the plots and residuals, 
# am not selecting this model as the "final" model
vif(lm_7)


```



## Model with CO.GT.and PT08.S2.NMHC. 
 
```{r Air multiple regression 2 predictors CO.GT. and PT08.S2.NMHC.}

# seems to be quadratic, also built a model with PT08.S2.NMHC.^2, however did not improve the model

lm_8 <- lm(C6H6.GT. ~ CO.GT. + PT08.S2.NMHC., airq_norm)
summary(lm_8)
residuals(lm_8) %>% hist(main = "residuals multi regression CO.GT. + PT08.S2.NMHC. predictors")

plot(lm_8, which = 1)

# vif >5: indicates multicollinearity
vif(lm_8)
```

## Final model with the following predictors: CO.GT.and PT08.S1.CO.

```{r Air multiple regression 2 predictors CO.GT. and PT08.S1.CO.}

# predictors significant, best residuals with a few outliers on plot, 
# high R^2 (larger than single regression with CO) and low vif

lm_9 <- lm(C6H6.GT. ~ CO.GT. + PT08.S1.CO., airq_norm)
summary(lm_9)
residuals(lm_9) %>% hist(main = "residuals multi regression CO.GT. + PT08.S1.CO predictors")

plot(lm_9, which = 1)

# vif < 5: 
#indicates that predictors are not redunant (not providing overlapping data to inform response)
vif(lm_9)

```






