---
title: "HW_2_1"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(tidyr)
library(reshape2)
library(plyr)
```

Task1.
Q1. Scatter plot facetted by set
```{r message=FALSE, warning=FALSE}
anscombe <- as.data.frame(anscombe)
head(anscombe)
nrow(anscombe)
#generate levels to indicate which group each data point belong to 
levels <- gl(4, nrow(anscombe))
levels
#Group it in dataframe
anscombe_grouped <- with(anscombe, data.frame(x=c(x1,x2,x3,x4),y=c(y1,y2,y3,y4), set=levels))
anscombe_grouped
#Make scattterplots
ggplot(anscombe_grouped, aes(x,y))+
  geom_point(size=3) +
  facet_wrap(~set)
```
Q2. Summary calculation(mean, sd) grouped by set
```{r message=FALSE, warning=FALSE}
#Mean
aggregate(cbind(x, y) ~ set, anscombe_grouped, mean)
#SD
aggregate(cbind(x, y) ~ set, anscombe_grouped, sd)
```
Q3. Pearson's correlation by set and non-parametric, and p-value
```{r message=FALSE, warning=FALSE}


anscombe_grouped%>% group_by(set) %>% summarise(cor_pearson = cor.test(x,y, method = 'pearson')$estimate, 
                                      cor_kendall = cor.test(x,y, method = 'kendall')$estimate,
                                      cor_spearmen = cor.test(x,y, method = 'spearman')$estimate)

anscombe_grouped %>% group_by(set) %>% summarise(p_pearson = cor.test(x,y, method = 'pearson')$p.value,
                                     p_kendall = cor.test(x,y, method = 'kendall')$p.value,
                                     p_spearmen = cor.test(x,y, method = 'spearman')$p.value)
```
Q4. Add geom smooth() to the plot
```{r message=FALSE, warning=FALSE}
ggplot( anscombe_grouped, aes(x, y)) +
  geom_point() + 
  geom_smooth(method = lm, se = F) +
  facet_wrap(~ set) 

```
Task2.
Q1. Explore data set, clean if needed
```{r message=FALSE, warning=FALSE}
aq <- read.csv2("/Users/Lisa/Downloads/AirQualityUCI/AirQualityUCI.csv", header =T)
str(aq)
#We see some NA. Let's remove them
colSums(is.na(aq))
aq_cleaned <-  aq  %>% select_if(~sum(!is.na(.)) > 0)  %>% drop_na()
head(aq_cleaned)
#Delete strange variables that contain value 200
airq_fil <- aq_cleaned %>% filter_all(all_vars(. != -200))
#Convert to long dataset
air_long <- melt(airq_fil)
head(air_long)

```
Q2.Explore each variable independently 
```{r message=FALSE, warning=FALSE}

  ggplot(air_long, aes(value)) +
  geom_boxplot() +
  facet_wrap(~variable, scales="free")

 ggplot(air_long, aes(value)) +
  geom_histogram() +
  facet_wrap(~variable, scales="free")

```
Q3.Cross correlation
```{r message=FALSE, warning=FALSE}
correl <- as.matrix(airq_fil[,c(3:15)])
cor(correl, use="complete.obs", method="spearman")
heatmap(cor(correl, use="complete.obs", method="spearman"))
```
Q4. Response C6H6(GT)
```{r message=FALSE, warning=FALSE}
airq_fil %>%
  lm(data = ., airq_fil$C6H6.GT. ~ airq_fil$CO.GT.) %>%
  summary()
```
Q5. Build simple linear models with each predictor, check assumptions, response C6H6
```{r}
airq_fil %>%
  ggplot(aes(x= CO.GT., y= C6H6.GT.)) +
  geom_point() +
  geom_smooth(method="lm")
```
```{r}
airq_fil  %>%
  lm(data= .,C6H6.GT. ~ C6H6.GT.)%>%
  summary()
```

```{r}
#PT08.S1.CO
airq_fil %>%
  ggplot(aes(x= PT08.S1.CO., y= C6H6.GT.)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
airq_fil  %>%
  lm(data= .,C6H6.GT. ~ PT08.S1.CO.)%>%
  summary()
```
```{r}
#PT08.S2.NMHC
airq_fil %>%
  ggplot(aes(x= PT08.S2.NMHC., y= C6H6.GT.)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
airq_fil  %>%
  lm(data= .,C6H6.GT. ~ PT08.S2.NMHC.)%>%
  summary()

```

```{r}
#PT08.S3.NOx
airq_fil %>%
  ggplot(aes(x= PT08.S3.NOx., y= C6H6.GT.)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
airq_fil  %>%
  lm(data= .,C6H6.GT. ~ PT08.S3.NOx.)%>%
  summary()
```

```{r}
#PT08.S4.NO2
airq_fil %>%
  ggplot(aes(x= PT08.S4.NO2., y= C6H6.GT.)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
airq_fil  %>%
  lm(data= .,C6H6.GT. ~ PT08.S4.NO2.)%>%
  summary()
```

```{r}
#PT08.S5.O3
airq_fil %>%
  ggplot(aes(x= PT08.S5.O3., y= C6H6.GT.)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
airq_fil  %>%
  lm(data= .,C6H6.GT. ~ PT08.S5.O3.)%>%
  summary()
```

```{r}
# NMHC.GT
airq_fil %>%
  ggplot(aes(x= NMHC.GT., y= C6H6.GT.)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
airq_fil  %>%
  lm(data= .,C6H6.GT. ~ NMHC.GT.)%>%
  summary()
```

```{r}
#NOx.GT
airq_fil %>%
  ggplot(aes(x= NOx.GT., y= C6H6.GT.)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
airq_fil  %>%
  lm(data= .,C6H6.GT. ~ NOx.GT.)%>%
  summary()
```


```{r}
#NO2.GT
airq_fil %>%
  ggplot(aes(x= NO2.GT., y= C6H6.GT.)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
airq_fil  %>%
  lm(data= .,C6H6.GT. ~ NO2.GT.)%>%
  summary()
```

```{r}
#T
airq_fil %>%
  ggplot(aes(x= T, y= C6H6.GT.)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
airq_fil  %>%
  lm(data= .,C6H6.GT. ~ T)%>%
  summary()
```

```{r}
#RH
airq_fil %>%
  ggplot(aes(x= RH, y= C6H6.GT.)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
airq_fil  %>%
  lm(data= .,C6H6.GT. ~ RH)%>%
  summary()

```
```{r}
airq_fil %>%
  ggplot(aes(x= AH, y= C6H6.GT.)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
airq_fil  %>%
  lm(data= .,C6H6.GT. ~ AH)%>%
  summary()
```
Q6. For 2-3 of the models create train-test sets, plot the model for the test set color real and predicted points differently; R^2 and p-value to title 

```{r}
#PT08.S5.O3
set.seed(42)
#Separated data in 75:25%
sample <- sample.int(n = nrow(airq_fil), size = floor(.75*nrow(airq_fil)))
training <- airq_fil[sample, ]
test <- airq_fil[-sample, ]
train_PT <- training[,c("C6H6.GT.", "PT08.S5.O3." )] 
test_PT <- test[,c("C6H6.GT.", "PT08.S5.O3." )]
model_PT <- lm(data = train_PT , C6H6.GT. ~ PT08.S5.O3.)
#summary
summary <- summary(model_PT)
print(summary)
#Predicition
pred_PT <- predict(model_PT, newdata = test_PT) 
test$PT08.S5.O3._pred <- pred_PT
#make combined dataset
train_PT$airq_fil <- "train"
test_PT$airq_fil <- "test" 
test_PT[1:(nrow(test_PT)/2),3] <- "real"

all_PT <- rbind(train_PT, test_PT)
```

```{r}
ggplot(data = all_PT, aes(x = PT08.S5.O3.,
y = C6H6.GT., 
color=airq_fil)) +
  geom_point() +
geom_smooth(method = "lm", color = "black") +
ggtitle(paste("R2", round(summary$r.squared, 3), sep = ": "),
paste("pvalue <2e-16" ))
```


```{r}
# NMHC.GT
set.seed(42)
#Separated data in 75:25%
sample <- sample.int(n = nrow(airq_fil), size = floor(.75*nrow(airq_fil)))
training <- airq_fil[sample, ]
test <- airq_fil[-sample, ]
train_MH <- training[,c("C6H6.GT.", "NMHC.GT." )] 
test_MH <- test[,c("C6H6.GT.", "NMHC.GT." )]
model_MH <- lm(data = train_MH , C6H6.GT. ~ NMHC.GT.)
#summary
summary <- summary(model_MH)
print(summary)

#Predicition
pred_MH <- predict(model_MH, newdata = test_MH) 
test$NMHC.GT._pred <- pred_MH
#make combined dataset
train_MH$airq_fil <- "train"
test_MH$airq_fil <- "test" 
test_MH[1:(nrow(test_MH)/2),3] <- "real"

all_PT <- rbind(train_MH, test_MH)
#trained and test should be similar shape
```

```{r}
ggplot(data = all_PT, aes(x = NMHC.GT.,
y = C6H6.GT., 
color=airq_fil)) +
  geom_point() +
geom_smooth(method = "lm", color = "black") +
ggtitle(paste("R2", round(summary$r.squared, 3), sep = ": "),
paste("pvalue <2e-16" ))

```





