---
title: "Statistics in R: Task 1"
author: "Liuaza Etezova"
output: pdf_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/Nexenta/prg/r/')
```


```{r packages, message=F}
library(ggplot2)
library(dplyr)
library(tidyr)
library(chron)
library(corrplot)
library(ggpubr)
```

# Anscombe

```{r anscombe}
data(anscombe)
head(anscombe)
```

```{r anscombeL}
anscombeL <- data.frame(
    set = rep(1:4, each=11),
    x = unlist(anscombe[,c(1:4)]),
    y = unlist(anscombe[,c(5:8)]))
rownames(anscombeL) <- NULL
head(anscombeL)
```

## Scatter plot facetted by set
```{r scatter plot}
ggplot(data=anscombeL, mapping=aes(x, y)) +
    geom_point() +
    facet_wrap(~set)
```

## Summary calculation (mean, sd) grouped by set
```{r groups}
by_set <- anscombeL %>%
  group_by(set)
```

```{r summary}
by_set %>%
  summarise_all(list(~mean(.), ~sd(.)))
```

## Correlations by set
```{r correlations}
by_set %>%
  summarise(cor_pearson = cor(x, y, method="pearson"), 
            cor_kendall = cor(x, y, method="kendall"), 
            cor_spearman = cor(x, y, method="spearman"))
```

```{r p-values, warning=F}
by_set %>%
  summarise(p_pearson = cor.test(x, y, method="pearson")$p.value, 
            p_kendall = cor.test(x, y, method="kendall")$p.value, 
            p_spearman = cor.test(x, y, method="spearman")$p.value)
```

## Scatter plot facetted by set with geom_smooth()
```{r geom_smooth}
ggplot(data=anscombeL, mapping=aes(x, y)) +
    geom_point() +
    geom_smooth(method='lm', se=F) +
    facet_wrap(~set)
```


# Air quality

```{r air quality}
air_quality <- read.csv('AirQualityUCI.csv', sep=';')
str(air_quality)
```

## Tidying
```{r tidying}
air_quality <- air_quality[-c(9358:nrow(air_quality)), 
                           !(names(air_quality) %in% c('X','X.1'))] 

air_quality$Date <- as.Date(as.character(air_quality$Date), "%d/%m/%Y")
air_quality$Time <- chron(times=air_quality$Time, format=c('h.m.s'))

columns <- c('CO.GT.', 'C6H6.GT.', 'T', 'RH', 'AH')
air_quality[columns] <- lapply(air_quality[columns], function(x) gsub(",", ".", x))
air_quality[columns] <- lapply(air_quality[columns], as.numeric)

str(air_quality)
```

## Checking each variable
```{r hist before}
airq_for_hist <- pivot_longer(air_quality, cols=CO.GT.:AH)
ggplot(airq_for_hist, aes(value)) + 
    geom_histogram() + 
    facet_wrap(~name, scales='free')
```

```{r hist scale, warning=F}
ggplot(airq_for_hist, aes(value)) + 
    geom_histogram() + 
    facet_wrap(~name, scales='free_y') +
    xlim(-300, 0)
```
-200 like the default value for this data. I will not delete all rows with this value because 
this way I will loose too much data (8,530 rows out of 9,357), but I will replace all individual 
values with NAs to pairwise exclude them in correlation analysis ("use='pairwise.complete.obs'" 
argument in "corr" function).

## Cleaning
```{r na}
air_quality[air_quality == -200] <- NA
```

```{r hist after, warning=F}
airq_for_hist <- pivot_longer(air_quality, cols=CO.GT.:AH)
ggplot(airq_for_hist, aes(value)) + 
    geom_histogram() + 
    facet_wrap(~name, scales='free')
```

## Cross correlations
```{r cross correlations}
cors <- cor(air_quality[3:15], use='pairwise.complete.obs', method='spearman')
```

```{r corrplot}
corrplot.mixed(cors, order='hclust', lower='number', tl.pos='lt', tl.col='black', 
               number.cex=0.55)
```

## Linear models, response - C6H6(GT)

### Checking the linearity of dependencies
```{r linearity, warning=F}
airq_for_lm <- pivot_longer(air_quality, cols=-c('Date','Time', 'C6H6.GT.'))
ggplot(airq_for_lm, aes(x=value, y=C6H6.GT.)) + 
    geom_point() + 
    geom_smooth(method='lm', se=F) + 
    facet_wrap(~name, scales='free')
```

### Linearizing
```{r linearizing plots, warning=F, message=F}
plot <- function(equation) {
  ggplot(air_quality, aes_string(x=equation, y='C6H6.GT.')) +
    geom_point() +
    geom_smooth(method='lm', se=F)
}

equations <- c('NMHC.GT.', '(NMHC.GT.)^(0.7)',
               'PT08.S1.CO.', '(PT08.S1.CO.)^2',
               'PT08.S2.NMHC.', '(PT08.S2.NMHC.)^2',
               'PT08.S3.NOx.', '1/(PT08.S3.NOx.)^2',
               'PT08.S4.NO2.', '(PT08.S4.NO2.)^(2.5)',
               'PT08.S5.O3.', '(PT08.S5.O3.)^(1.5)')
plots <- lapply(equations, plot)
ggarrange(plotlist=plots, ncol=2)
```

```{r linearizing}
airq_l <- air_quality
airq_l$`NMHC.GT.^0.7` <- air_quality$NMHC.GT.^(0.7)
airq_l$`PT08.S1.CO.^2` <- air_quality$PT08.S1.CO.^2
airq_l$`PT08.S2.NMHC.^2` <- air_quality$PT08.S2.NMHC.^2
airq_l$`PT08.S3.NOx.^2` <- 1/air_quality$PT08.S3.NOx.^2
airq_l$`PT08.S4.NO2.^2.5` <- air_quality$PT08.S4.NO2.^(2.5)
airq_l$`PT08.S5.O3.^1.5` <- air_quality$PT08.S5.O3.^(1.5)
```

### Building linear models
```{r lm}
benzene_lm_summary <- function(predictor) {
  airq_l %>%
    lm(data = ., C6H6.GT. ~ predictor) %>%
    summary()
}

lapply(airq_l[, !(names(airq_l) %in% c('Date', 'Time', 'C6H6.GT.', 
                                       'T', 'RH', 'AH', 
                                       'NMHC.GT.', 'PT08.S1.CO.', 'PT08.S2.NMHC.', 
                                       'PT08.S3.NOx.', 'PT08.S4.NO2.', 'PT08.S5.O3.'))], 
       benzene_lm_summary)
```

### Selected predictors
PT08.S2(NMHC)^2 - titanium dioxide (R^2 = 0.9996),
CO(GT) - carbon monoxide (R^2 = 0.8669), 
NMHC(GT)^0.7 - nonmethane hydrocarbons (R^2 = 0.8486).
```{r}
set.seed(42)
sample <- sample.int(n=nrow(airq_l), size=floor(.75*nrow(airq_l)))
train <- airq_l[sample, ]
test <- airq_l[-sample, ]

model_PT08S2 <- lm(data=train, C6H6.GT. ~ `PT08.S2.NMHC.^2`)
model_CO <- lm(data=train, C6H6.GT. ~ CO.GT.)
model_NMHC <- lm(data=train, C6H6.GT. ~ `NMHC.GT.^0.7`)

test$C6H6byPT08S2 <- predict(model_PT08S2, newdata=test)
test$C6H6byCO <- predict(model_CO, newdata=test)
test$C6H6byNMHC <- predict(model_NMHC, newdata=test)
```

### Plots
```{r, warning=F}
plot <- function(lm_info) {
    ggplot() +
    geom_point(data=train, aes_string(x=lm_info[[1]], y='C6H6.GT.', 
                                      color=shQuote('train'), shape=shQuote('train'))) +
    geom_point(data=test, aes_string(x=lm_info[[1]], y='C6H6.GT.', 
                                     color=shQuote('test'), shape=shQuote('test'))) + 
    geom_smooth(method='lm', se=F, data=train, aes_string(x=lm_info[[1]], y='C6H6.GT.'), 
                color='black') + 
    geom_point(data=test, aes_string(x=lm_info[[1]], y=lm_info[[2]], 
                                     color=shQuote('predicted'), 
                                     shape=shQuote('predicted'))) + 
    scale_colour_manual(name="data", breaks=c('train', 'test', 'predicted'), 
                        values=c(train="blue", test="red", predicted="green3")) + 
    scale_shape_manual(name="data", breaks=c('train', 'test', 'predicted'), 
                       values=c(train=1, test=2, predicted=20)) + 
    ggtitle(paste("R-squared = ", lm_info[[3]]$adj.r.squared, '\n', 
                  "p-value = ", lm_info[[3]]$coefficients[,4][[2]]))
}

# list(predictor, predicted, linear model summary)
PT08S2 <- list('PT08.S2.NMHC.^2', 'C6H6byPT08S2', summary(model_PT08S2))
CO     <- list('CO.GT.', 'C6H6byCO', summary(model_CO))
NMHC   <- list('NMHC.GT.^0.7', 'C6H6byNMHC', summary(model_NMHC))

info <- list(PT08S2 = PT08S2, CO = CO, NMHC = NMHC)
lapply(info, plot)
```
