---
title: "Task_2.1"
author: "NataliaBaymacheva"
date: "5/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(corrplot)
```

Look at the data
```{r}
df <- anscombe
head(df)
```

Converting the data to a better format
```{r}
df1 <- data.frame(cbind(df$x1, df$y1, 'set' = 1))
df2 <- data.frame(cbind(df$x2, df$y2, 'set' = 2))
df3 <- data.frame(cbind(df$x3, df$y3, 'set' = 3))
df4 <- data.frame(cbind(df$x4, df$y4, 'set' = 4))
df <- rbind(df1, df2, df3, df4)
names(df)[names(df) == 'V1'] <- 'x'
names(df)[names(df) == 'V2'] <- 'y'

head(df)
```

Plotting the data
```{r, message=FALSE, warning=FALSE}
df %>%
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  facet_wrap(~set) +
  geom_smooth()
```

Mean/sd summary table
```{r}
df %>%
  group_by(set) %>%
  summarise(mean_x=mean(x), sd_x = sd(x), mean_y=mean(y), sd_y = sd(y))
```

Pearson's/non-parametric summary table
```{r, warning=FALSE}
df %>%
  group_by(set) %>%
  summarise(pearson_est=cor.test(x, y)$estimate, pearson_p=cor.test(x, y)$p.value,
            kendall_est=cor.test(x, y, method = 'kendall')$estimate, kendall_p=cor.test(x, y, method = 'kendall')$p.value,
            spearman_est=cor.test(x, y, method = 'spearman')$estimate, spearman_p=cor.test(x, y, method = 'spearman')$p.value)
```

```{r}
rm(list = ls())
```

AIRQUALITY DATA

Loading libraries
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(corrplot)
```

```{r}
airquality_data <- read.csv2('~/R/AirQualityUCI.csv', sep = ';', blank.lines.skip = T)[,1:15]
head(airquality_data)
```

There are many NA lines, get rid of them
```{r}
airquality_data <- drop_na(airquality_data)
```

Transforming dates to date types
```{r}
airquality_data$Date <- as.Date(airquality_data$Date)
```

Checcking the structures of the cols
```{r}
str(airquality_data)
```

Change all -200 to na
```{r}
for (col in 3:15) {
  airquality_data[which(airquality_data[col] == -200), col] <- NA
}
```

Check distributions
```{r}
airquality_long <- pivot_longer(airquality_data, cols = 3:15, names_to = 'measure', values_to = 'value', values_drop_na = T)
airquality_long %>%
  group_by(measure) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~measure, scales = 'free')
```

Not all data is normalli distributed and all variables have different scaling

Perform data normalization

```{r}
airquality_norm_long <- airquality_long
airquality_norm_long$value <- log10(airquality_norm_long$value)
```

```{r}
airqual_norm <- airquality_data[, 3:15]
airqual_norm <- log10(airqual_norm) 
```

Plot normilized data
```{r}
airquality_norm_long %>%
  group_by(measure) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~measure, scales = 'free')
```

Plotting each pair of variables
```{r}
pairs(airqual_norm)
```

Cross-correlation
```{r}
corr_kendall <- cor(airquality_data[, 3:15], use = 'complete.obs', method = 'kendall')
corrplot(corr_kendall)
```

```{r}
corr_spearman <- cor(airquality_data[, 3:15], use = 'complete.obs', method = 'spearman')
corrplot(corr_spearman)
```

Looks like spearman method detects stronger correlation between benzene and other variables.

Plotting Response of benzene for each predictor

``` {r}
airqual_norm %>%
  ggplot(aes(x = CO.GT., y = C6H6.GT.)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
airqual_norm %>%
  ggplot(aes(x = PT08.S1.CO., y = C6H6.GT.)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
airqual_norm %>%
  ggplot(aes(x = NMHC.GT., y = C6H6.GT.)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
airqual_norm %>%
  ggplot(aes(x = PT08.S2.NMHC., y = C6H6.GT.)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
airqual_norm %>%
  ggplot(aes(x = NOx.GT., y = C6H6.GT.)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
airqual_norm %>%
  ggplot(aes(x = PT08.S3.NOx., y = C6H6.GT.)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
airqual_norm %>%
  ggplot(aes(x = NO2.GT., y = C6H6.GT.)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
airqual_norm %>%
  ggplot(aes(x = PT08.S4.NO2., y = C6H6.GT.)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
airqual_norm %>%
  ggplot(aes(x = PT08.S5.O3., y = C6H6.GT.)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
set.seed(0)
```

Creating train/test data
```{r}
aq_sample <- sample.int(n = nrow(airqual_norm), size = floor(.75*nrow(airqual_norm)))
train_air_qual <- airqual_norm[aq_sample, ]
test_air_qual <- airqual_norm[-aq_sample, ]
```

Building models
```{r}
l_m_CO <- lm(data=train_air_qual, C6H6.GT.~CO.GT.)
CO_sum <- summary(l_m_CO)
R_CO <- round(CO_sum$r.squared, 4)
p_CO <- round(CO_sum$coefficients[2, 4], 4)
title_CO <- paste('R^2 =', as.character(R_CO), 'p-value =', as.character(p_CO))
```

```{r}
l_m_NO <- lm(data=train_air_qual, C6H6.GT.~NOx.GT.)
NO_sum <- summary(l_m_NO)
R_NO <- round(NO_sum$r.squared, 4)
p_NO <- round(NO_sum$coefficients[2, 4], 4)
title_NO <- paste('R^2 =', as.character(R_NO), 'p-value =', as.character(p_NO))
```

```{r}
l_m_NO2 <- lm(data=train_air_qual, C6H6.GT.~NO2.GT.)
NO2_sum <-summary(l_m_NO2)
R_NO2 <- round(NO2_sum$r.squared, 4)
p_NO2 <- round(NO2_sum$coefficients[2, 4], 4)
title_NO2 <- paste('R^2 =', as.character(R_NO2), 'p-value =', as.character(p_NO2))
```

Prediction of the test data
```{r}
pred_CO <- predict(l_m_CO, newdata = test_air_qual)
test_air_qual$pred_by_CO <- pred_CO

pred_NO <- predict(l_m_NO, newdata = test_air_qual)
test_air_qual$pred_by_NO <- pred_NO

pred_NO2 <- predict(l_m_NO2, newdata = test_air_qual)
test_air_qual$pred_by_NO2 <- pred_NO2

test <- test_air_qual %>%
  select(C6H6.GT., pred_by_CO, CO.GT., pred_by_NO, NOx.GT., pred_by_NO2, NO2.GT.) %>%
  drop_na()

head(test)
```

Building plots

To plot graph with predict and test data I only chose 30 random values. (If plot all predicted values they cover over linear regression line)
```{r}
test_for_graph <- test[sample(rownames(test), size = 30), ]
```

```{r}
ggplot() +
  geom_point(data = airqual_norm, aes(x = CO.GT., y = C6H6.GT.)) +
  geom_smooth(data = airqual_norm, method = 'lm', aes(CO.GT., C6H6.GT.)) +
  geom_point(data = test_for_graph, aes(x = CO.GT., y = pred_by_CO), color = 'red') +
  geom_point(data = test_for_graph, aes(x = CO.GT., y = C6H6.GT.), color = 'green') +
  labs(title = title_CO)
```

```{r}
ggplot() +
  geom_point(data = airqual_norm, aes(x = NOx.GT., y = C6H6.GT.)) +
  geom_smooth(data = airqual_norm, method = 'lm', aes(NOx.GT., C6H6.GT.)) +
  geom_point(data = test_for_graph, aes(x = NOx.GT., y = pred_by_NO), color = 'red') +
  geom_point(data = test_for_graph, aes(x = NOx.GT., y = C6H6.GT.), color = 'green') +
  labs(title = title_NO)
```

```{r}
ggplot() +
  geom_point(data = airqual_norm, aes(x = NO2.GT., y = C6H6.GT.)) +
  geom_smooth(data = airqual_norm, method = 'lm', aes(NO2.GT., C6H6.GT.)) +
  geom_point(data = test_for_graph, aes(x = NO2.GT., y = pred_by_NO2), color = 'red') +
  geom_point(data = test_for_graph, aes(x = NO2.GT., y = C6H6.GT.), color = 'green') +
  labs(title = title_NO2)
```