---
title: "Task 22"
author: "Lisa Skalon"
date: "5/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE)
```


```{r results="hide"}
library('ggplot2')
library('ggpubr')
library('dplyr')
library('tidyr')
library('broom')
library('lubridate')
library('reshape2')
library('car')

```


## Airquality - part 2

We read df again and clean it almost the same as in the part 1. 
```{r}
# read
df <- read.csv("./AirQualityUCI.csv", stringsAsFactors = FALSE, sep = ';')

# examine the data structure
str(df)

# convert date and time to special type
df$date_time = dmy_hms(paste(df$Date, df$Time))

# remove unuseful cols
df <- df[,c(3:15,18) ]

# chr to numeric
char_columns <- sapply(df, is.character)
df[ , char_columns] <- lapply(df[ , char_columns] , function(x) as.numeric(gsub(",", ".", x)))

# remove na
df <- na.omit(df)

# value -200 is suspicious
# let values -200 be NA
df_no200 <- df[, 1:13]
df_no200[] <- sapply(df[, 1:13] , function(x) {x[grep("-200", x)] = NA; return((x))}) 
str(df_no200)

summary(df_no200)

# We remove almost all suspicious values, but the column NMHC.GT. contains too many of them, 
# so for that column we replace -200 with median, because we don`t want to loose information
df_no200 <- df_no200[!(is.na(df_no200$AH) | is.na(df_no200$NO2.GT.) | is.na(df_no200$CO.GT.)),]

df_narm <- replace(df_no200, TRUE, lapply(df_no200, function(x) replace(x, is.na(x), median(x, na.rm = TRUE))))

summary(df_narm)

```


For exploring collinearity of predictors we can simple check for cross-correlations between 
variables. If the correlation is greater than 0.8/lower than -0.8, probably these predictors
depend on each other and should be removed from the analysis.

```{r}
# correlation heatmap - more comportable way to find cross-correlations
cor_mtx <- (cor(df_narm))
cor_mtx

# heatmap
ggplot(data = melt(cor_mtx), aes(Var2, Var1, fill = value))+
 geom_tile(color = "black")+
 theme(axis.text.x = element_text(angle = 90))+
  coord_fixed()
```

Then 3 or more predictors depend on each other, we are talking about multicollinearity. 
It can be measured by VIF score -the variance inflation factor, which evaluate how much the variance of a regression coefficient is inflated due to multicollinearity in the model. 
We can calculate VIF for each predictor:

```{r}
fit <- lm(data = df_narm, formula =  C6H6.GT. ~ .-NMHC.GT. ) 
summary(fit)

vif(fit)

```

It seeams that many predictors have too big VIF score (>10). These predictors shouldn`t be
included into our model.


Now we can choose predictors which are significant and correlated with our response 
but have low VIF score. Let`s try out some of them.


First of all, we check the most significant predictors CO.GT. and  PT08.S2.NMHC.
```{r}
fit1 <- lm(data = df_narm, formula =  C6H6.GT. ~  CO.GT.  +  PT08.S2.NMHC.) 
summary(fit1)

vif(fit1)

```
VIF score for them is not very big, but quantiles are not symmetrical. 
We can check the residuals of this model: residuals represent variation left unexplained by the model.

```{r}
plot(fit1, which=1, col=c("blue"))
plot(fit1, which=2, col=c("red"))

```
The residuals doesn`t look linear. The are not normally distributed. The model is not good enough. 


We will try to choose either CO.GT. or  PT08.S2.NMHC. as predictors because they both are 
very significant but strongly correlated. 

Let`s try PT08.S5.O3. and RH  
```{r}
fit2 <- lm(data = df_narm, formula =  C6H6.GT. ~  CO.GT.  +  PT08.S5.O3. + RH) 
summary(fit2)

vif(fit2)

```


```{r}
plot(fit2, which=1, col=c("blue"))
plot(fit2, which=2, col=c("red"))

```

Quantiles are symmetrical, VIF is ok, residuals show linearity and are almost normally distributed. 


```{r}
fit3 <- lm(data = df_narm, formula =  C6H6.GT. ~  PT08.S5.O3.+ PT08.S3.NOx.) 
summary(fit3)

vif(fit3)

```


```{r}
plot(fit3, which=1, col=c("blue"))
plot(fit3, which=2, col=c("red"))

```
We have some problems with residuals here.


```{r}
fit4 <- lm(data = df_narm, formula =  C6H6.GT. ~ CO.GT. + PT08.S5.O3. + NO2.GT. ) 
summary(fit4)
vif(fit4)

```

```{r}
plot(fit4, which=1, col=c("blue"))
plot(fit4, which=2, col=c("red"))

```
Residuals look ok, quantiles are ok, this model is the best one.

So, we will choose model 4: C6H6.GT. ~ CO.GT. + PT08.S5.O3. + NO2.GT.


Should we transform our data somehow? Let`s check the distribution of variables.

```{r}
df_narm%>%
  mutate(id=c(1:nrow(df_narm)))%>%
  melt(measure.vars=c(1:13))->long

ggplot(long, aes(value)) + 
  geom_histogram(bins=40) + 
  facet_wrap(~variable, scales = "free")

```
Not all of the variables look normal. Let`s try log transformation.

```{r}
fitlog <- lm(data = df_narm, formula =  log(C6H6.GT.) ~ log(CO.GT.) + log(PT08.S5.O3.) + log(NO2.GT.) ) 
summary(fitlog)
vif(fitlog)

```

```{r}
plot(fitlog, which=1, col=c("blue"))
plot(fitlog, which=2, col=c("red"))

```

Residuals look a bit skewed.

# So, we choose model 4 as the final model (C6H6.GT. ~ CO.GT. + PT08.S5.O3. + NO2.GT.)
