---
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries
```{r, include=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(reshape2)
library(corrplot)
```



##Air quality data set - Multiple linear regreession 
Data Set Information:

The dataset contains 9358 instances of hourly averaged responses from an array of 5 metal oxide chemical sensors embedded in an Air Quality Chemical Multisensor Device. The device was located on the field in a significantly polluted area, at road level,within an Italian city. Data were recorded from March 2004 to February 2005 (one year)representing the longest freely available recordings of on field deployed air quality chemical sensor devices responses. Ground Truth hourly averaged concentrations for CO, Non Metanic Hydrocarbons, Benzene, Total Nitrogen Oxides (NOx) and Nitrogen Dioxide (NO2) and were provided by a co-located reference certified analyzer. Evidences of cross-sensitivities as well as both concept and sensor drifts are present as described in De Vito et al., Sens. And Act. B, Vol. 129,2,2008 (citation required) eventually affecting sensors concentration estimation capabilities. Missing values are tagged with -200 value.
This dataset can be used exclusively for research purposes. Commercial purposes are fully excluded.

Attribute Information:

0 Date (DD/MM/YYYY)
1 Time (HH.MM.SS)
2 True hourly averaged concentration CO in mg/m^3 (reference analyzer)
3 PT08.S1 (tin oxide) hourly averaged sensor response (nominally CO targeted)
4 True hourly averaged overall Non Metanic HydroCarbons concentration in microg/m^3 (reference analyzer)
5 True hourly averaged Benzene concentration in microg/m^3 (reference analyzer)
6 PT08.S2 (titania) hourly averaged sensor response (nominally NMHC targeted)
7 True hourly averaged NOx concentration in ppb (reference analyzer)
8 PT08.S3 (tungsten oxide) hourly averaged sensor response (nominally NOx targeted)
9 True hourly averaged NO2 concentration in microg/m^3 (reference analyzer)
10 PT08.S4 (tungsten oxide) hourly averaged sensor response (nominally NO2 targeted)
11 PT08.S5 (indium oxide) hourly averaged sensor response (nominally O3 targeted)
12 Temperature in °C
13 Relative Humidity (%)
14 AH Absolute Humidity

Loading csv file
```{r}
AQ_raw <- read.csv("/home/aleksandar/Desktop/R statistics 2/hw_1_Rstat2/AirQualityUCI/AirQualityUCI.csv", sep = ";")

```

Removing NAs
Since missing values are denoted with -200 in this data set, first step would be to change all -200 values to NA and then remove NAs

```{r}
AQ_NAs <- AQ_raw %>% na_if(-200) %>% na_if("-200,0")
#removing empty columns
toDrop <- c("X", "X.1")
AQ_NAs <- AQ_NAs[,!(names(AQ_NAs)) %in% toDrop]

```

Exploring NAs, after visually checking data set it seems most NAs are from NMHC.GT column
```{r}
sapply(AQ_NAs, function(x) sum(is.na(x)))
```
Since NMHC.GT has a lot of NAs using complete cases would drop about 90% of cases. I think its a good option to remove NMHC.GT from the main analyses.


```{r}
toDrop <- c("NMHC.GT.", "Date", "Time") #also dropped Date and Time colums since I wont be using them
AQ_NAs <- AQ_NAs[,!(names(AQ_NAs)) %in% toDrop]
complete_AQrows <- complete.cases(AQ_NAs)
AQ <- AQ_NAs[complete_AQrows,]
```


Checking for non numeric variables
```{r}
sapply(AQ, function(x)is.numeric(x))

AQnum <- data.frame(sapply(AQ, function(x) as.double(gsub(",",".",x)))) #converting to double

sapply(AQnum, function(x)is.numeric(x))
```

Exploring the data individually 
```{r}
AQlong <- gather(AQnum)

ggplot(AQlong, aes(value)) + geom_histogram() + facet_wrap(~key, scales = "free")
  
```

```{r}
ggplot(AQlong, aes(value)) + geom_boxplot() + facet_wrap(~key, scales = "free")

```

Finding pairwise correlations for data set
Response for further analyses is C6H6.GT.

Multicollinearity
```{r}
corrplot.mixed(corrplot(cor(AQnum)), lower.col = "black", number.cex = .7)


```



##Visualising and inspecting some pairs of columns more closely

```{r}
d <- melt(AQnum, id.vars = "C6H6.GT.")
ggplot(d, aes(x = value, y = C6H6.GT.)) + geom_point() + geom_smooth(method = lm) + facet_wrap(~variable, scales = "free")

```

Before constructing a model I decided to look into NMHC vs C6H6, there seems to be very little devieation here, but I suspect that the curve of the scatter plot isnt suited for a linear model.
```{r}
ggplot(AQnum, aes(x = PT08.S2.NMHC., y = C6H6.GT.)) + geom_point() + geom_smooth(method = lm)
```

I tried squaring the NMCH column  to se if it would improve result
```{r}
AQnum$PT08.S2.NMHC. <- AQ$PT08.S2.NMHC.^2
```

Not it is almost a perfect fit
```{r}
ggplot(AQnum, aes(x = PT08.S2.NMHC., y = C6H6.GT.)) + geom_point() + geom_smooth(method = lm)

```


Multiple regression model - predictors: CO.GT. +  PT08.S1.CO. + PT08.S2.NMHC. + PT08.S4.NO2. + PT08.S5.O3. 
```{r}
ML_mod <- lm(formula = C6H6.GT. ~ CO.GT. +  PT08.S1.CO. + PT08.S2.NMHC. + PT08.S4.NO2. + PT08.S5.O3. , data = AQnum)
summary(ML_mod)
```
PT08.S1.CO. has lowered pvalue when included in multiple regression model in contrast when being the sole predictor. I will remove it from the model since  0.0908 > 0.05

```{r}
ML_mod <- lm(formula = C6H6.GT. ~ CO.GT. + PT08.S2.NMHC. + PT08.S4.NO2. + PT08.S5.O3. , data = AQnum)
summary(ML_mod)
```


This doesnt look so good, residuals are not symetric
```{r}
boxplot(ML_mod$residuals)
```
Residuals vs fitted - how well do fitted value align to residuals - it means residuals are uneuqally spread around the data
```{r}
plot(ML_mod, which = 1)
```

QQplot - are theoretical quantiles the same as actual ones - this means that residuals at one point are not random and that there is something skewing the model
```{r}
plot(ML_mod, which = 2)

```

Try to build model again and repeat analysis, removing NMHC from model

```{r}
ML_mod <- lm(formula = C6H6.GT. ~ CO.GT. + PT08.S4.NO2. + PT08.S5.O3. , data = AQnum)
summary(ML_mod)
```

Now it looks acceptable
```{r}
plot(ML_mod, which = 1)
```

```{r}
plot(ML_mod, which = 2)

```


Perfoming train vs test data LM analysis
```{r}
set.seed(42)
sample <- sample.int(n = nrow(AQnum), size = floor(0.75 * nrow(AQnum)))
train <- AQnum[sample,]
test <- AQnum[-sample,]

new_modCO <-  lm(formula = C6H6.GT. ~ CO.GT. + PT08.S4.NO2. + PT08.S5.O3. , data = train)
summary(new_modCO)
```


Prediction using trained model

```{r}
pred <- predict(new_modCO, newdata = test)
test$CO.GT._predicted <- pred

```

Constructing plot
```{r}
ggplot() + geom_point(data = train, aes(x= CO.GT., y = C6H6.GT.)) + geom_smooth(data =train, aes(x=CO.GT., C6H6.GT.),method = lm) + geom_point(data = test,  colour = "red", aes(x= CO.GT., y = C6H6.GT.)) + geom_point(data = test,  colour = "green", aes(x= CO.GT. ,y = CO.GT._predicted)) + ggtitle("C6H6 vs CO, R^2: 0.865, p value: 2.2e-16")
```




