---
title: "Task 1"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(corrplot)
```


## Air Quality data set

```{r}
airq_data <- read.table('/home/marina/Загрузки/AirQualityUCI.csv', sep = ';', header = TRUE, dec = ",")

```

• Deleting NA columns and rows. Averaged concentration CO, Averaged Benzene concentration, Temperature, Humidity are factor columns in data, so it will be better to convert them into numeric. Actually, I don't need date and time for futher analysis, so I will slice the data.

```{r}
airq_data <- airq_data %>% select_if(~sum(!is.na(.)) > 0)  %>% drop_na()
airq_data[c('CO.GT.', 'C6H6.GT.', 'T', 'RH', 'AH')] <-  sapply(airq_data[c('CO.GT.', 'C6H6.GT.', 'T', 'RH', 'AH')], as.numeric)
airq_data <- airq_data[, c(3:15)]
```

• Replacing -200 to NA. 

```{r}
airq_data <- sapply(airq_data, function(x){ifelse (x == -200, NA, x)})
airq_data <- as.data.frame(airq_data)
```

• Exploring the variables (I make it in a 3 parts to better visualizing)

```{r}
ggplot(gather(airq_data[, c(1:4)], cols, value), aes(x = value)) + 
  geom_histogram(binwidth = 50) + facet_wrap(.~cols)
```

```{r}
ggplot(gather(airq_data[, c(5:8)], cols, value), aes(x = value)) + 
  geom_histogram(binwidth = 50) + facet_wrap(.~cols)

```

```{r}
ggplot(gather(airq_data[, c(9:13)], cols, value), aes(x = value)) + 
  geom_histogram(binwidth = 50) + facet_wrap(.~cols)
```

Ok, so it looks like a lot of variables have outliers and all of them in a different scale range
So, as for me,  it will be better to log-transform the data

• Log-transformation

```{r message=FALSE, warning=FALSE}
airq_data <- log10(airq_data)
```

So, let's look now as an example for the first 4 columns:

```{r}
ggplot(gather(airq_data[, c(1:4)], cols, value), aes(x = value)) + 
  geom_histogram(binwidth = 0.1) + facet_wrap(.~cols)
```

• Relationships between all variables

```{r}
pairs(airq_data)
```


```{r}
cor_data <- cor(airq_data, method = 'spearman', use = 'pairwise.complete.obs')
corrplot(cor_data, method = 'circle')
```



Example of C6.H6.GT.:

```{r message=FALSE, warning=FALSE}
ggplot(airq_data, aes(x = CO.GT., y = C6H6.GT. )) + 
  geom_point() + 
  geom_smooth(method = 'lm')
```

```{r}
model_CO.GT. <- airq_data  %>%
   lm(data = .,  C6H6.GT. ~ CO.GT.)
residuals(model_CO.GT.) %>% hist()
```


```{r}
residuals(model_CO.GT.) %>% boxplot()
```


```{r}
plot(model_CO.GT., which = c(1,2))
```
```{r}
summary(model_CO.GT.)
```

So I decided to take a PT08.S3.NOx. as a predictor, because for me it seems a little bit better.
I checked the assumptions for each variable and took 4 best:

1) PT08.S5.O3.

```{r message=FALSE, warning=FALSE}
ggplot(airq_data, aes(x = PT08.S5.O3., y = PT08.S3.NOx. )) + 
  geom_point() + 
  geom_smooth(method = 'lm')
```

```{r}
model_PT08.S5.O3. <- airq_data %>% lm(data = ., PT08.S3.NOx. ~ PT08.S5.O3., na.action = na.exclude)
residuals(model_PT08.S5.O3.) %>% hist()
```


```{r}
residuals(model_PT08.S5.O3.) %>% boxplot()
```


```{r}
plot(model_PT08.S5.O3., which = c(1,2))
```

```{r message=FALSE, warning=FALSE}
summary(model_PT08.S5.O3.)
```

Prediction:

```{r paged.print=FALSE}
test_subset_PT08.S5.O3. <-  airq_data[which(row.names(airq_data) %in% sample(row.names(airq_data), 25, replace = FALSE)), c(10,7)]
test_PT08.S5.O3. <- data.frame(PT08.S5.O3. = test_subset_PT08.S5.O3.$PT08.S5.O3.)
test_subset_PT08.S5.O3.$pred_PT08.S3.NOx. <- predict(model_PT08.S5.O3., newdata = test_PT08.S5.O3.)
colnames(test_subset_PT08.S5.O3.) <- c('real_PT08.S5.O3.', 'real_PT08.S3.NOx.', 'pred_PT08.S3.NOx.')
head(test_subset_PT08.S5.O3.)
```
      
```{r message=FALSE, warning=FALSE}
R <- round(summary(model_PT08.S5.O3.)$adj.r.squared, digits = 4)
p <- round(summary(model_PT08.S5.O3.)$coefficients[2,4], digits = 3)
titl <- paste('R^2 =', as.character(R),', p-val =', as.character(p))
ggplot() + 
  geom_point(data = airq_data, aes(PT08.S5.O3., PT08.S3.NOx.)) + 
  geom_smooth(data = airq_data, aes(PT08.S5.O3., PT08.S3.NOx.), method = 'lm') +
  geom_point(data = test_subset_PT08.S5.O3., aes(real_PT08.S5.O3., real_PT08.S3.NOx.), color = 'red') + 
  geom_point(data = test_subset_PT08.S5.O3., aes(real_PT08.S5.O3., pred_PT08.S3.NOx.), color = 'green') +
  labs(title = titl)
```

3) PT08.S2.NMHC.

```{r message=FALSE, warning=FALSE}
ggplot(airq_data, aes(x = PT08.S2.NMHC., y = PT08.S3.NOx. )) + 
  geom_point() + 
  geom_smooth(method = 'lm')
```

```{r}
model_PT08.S2.NMHC. <- airq_data  %>% lm(data = .,  PT08.S3.NOx. ~ PT08.S2.NMHC.)
residuals(model_PT08.S2.NMHC.) %>% hist()
```


```{r}
residuals(model_PT08.S2.NMHC.) %>% boxplot()
```


```{r}
plot(model_PT08.S2.NMHC., which = c(1,2))
```

```{r message=FALSE, warning=FALSE}
summary(model_PT08.S2.NMHC.)
```

Prediction:

```{r paged.print=FALSE}
test_subset_PT08.S2.NMHC. <-  airq_data[which(row.names(airq_data) %in% sample(row.names(airq_data), 25, replace = FALSE)), c(5,7)]
test_PT08.S2.NMHC. <- data.frame(PT08.S2.NMHC. = test_subset_PT08.S2.NMHC.$PT08.S2.NMHC.)
test_subset_PT08.S2.NMHC.$pred_PT08.S3.NOx. <- predict(model_PT08.S2.NMHC., newdata = test_PT08.S2.NMHC.)
colnames(test_subset_PT08.S2.NMHC.) <- c('real_PT08.S2.NMHC.', 'real_PT08.S3.NOx.', 'pred_PT08.S3.NOx.')
head(test_subset_PT08.S2.NMHC.)
```

```{r message=FALSE, warning=FALSE}
R <- round(summary(model_PT08.S2.NMHC.)$adj.r.squared, digits = 3)
p <- round(summary(model_PT08.S2.NMHC.)$coefficients[2,4], digits = 3)
titl <- paste('R^2 =', as.character(R),', p-val =', as.character(p))
ggplot() + 
  geom_point(data = airq_data, aes(PT08.S2.NMHC., PT08.S3.NOx.)) + 
  geom_smooth(data = airq_data, aes(PT08.S2.NMHC., PT08.S3.NOx.), method = 'lm') +
  geom_point(data = test_subset_PT08.S2.NMHC., aes(real_PT08.S2.NMHC., real_PT08.S3.NOx.), color = 'red') + 
  geom_point(data = test_subset_PT08.S2.NMHC., aes(real_PT08.S2.NMHC., pred_PT08.S3.NOx.), color = 'green') +
  labs(title = titl)
```

3) PT08.S1.CO.

```{r message=FALSE, warning=FALSE}
ggplot(airq_data, aes(x = PT08.S1.CO., y = PT08.S3.NOx. )) + 
  geom_point() + 
  geom_smooth(method = 'lm')
```

```{r}
model_PT08.S1.CO. <- airq_data  %>% lm(data = .,  PT08.S3.NOx. ~ PT08.S1.CO.)
residuals(model_PT08.S1.CO.) %>% hist()
```


```{r}
plot(model_PT08.S1.CO., which = c(1,2))
```

```{r message=FALSE, warning=FALSE}
summary(model_PT08.S1.CO.)
```

Prediction:

```{r paged.print=FALSE}
test_subset_PT08.S1.CO. <-  airq_data[which(row.names(airq_data) %in% sample(row.names(airq_data), 25, replace = FALSE)), c(2,7)]
test_PT08.S1.CO. <- data.frame(PT08.S1.CO. = test_subset_PT08.S1.CO.$PT08.S1.CO.)
test_subset_PT08.S1.CO.$pred_PT08.S3.NOx. <- predict(model_PT08.S1.CO., newdata = test_PT08.S1.CO.)
colnames(test_subset_PT08.S1.CO.) <- c('real_PT08.S1.CO.', 'real_PT08.S3.NOx.', 'pred_PT08.S3.NOx.')
head(test_subset_PT08.S1.CO.)
```

```{r message=FALSE, warning=FALSE}
R <- round(summary(model_PT08.S1.CO.)$adj.r.squared, digits = 3)
p <- round(summary(model_PT08.S1.CO.)$coefficients[2,4], digits = 3)
titl <- paste('R^2 =', as.character(R),', p-val =', as.character(p))
ggplot() + 
  geom_point(data = airq_data, aes(PT08.S1.CO., PT08.S3.NOx.)) + 
  geom_smooth(data = airq_data, aes(PT08.S1.CO., PT08.S3.NOx.), method = 'lm') +
  geom_point(data = test_subset_PT08.S1.CO., aes(real_PT08.S1.CO., real_PT08.S3.NOx.), color = 'red') + 
  geom_point(data = test_subset_PT08.S1.CO., aes(real_PT08.S1.CO., pred_PT08.S3.NOx.), color = 'green') +
  labs(title = titl)
```

4) NMHC.GT. - there are a lot of missing data, but it was just interesting that for the rest data the model and prediction is very good.
*But, of course we should not use this model, due to missing a lot of data.

```{r message=FALSE, warning=FALSE}
ggplot(airq_data, aes(x = NMHC.GT., y = PT08.S3.NOx.)) + 
  geom_point() + 
  geom_smooth(method = 'lm')
```

```{r}
model_NMHC.GT. <- airq_data %>%  lm(data = ., PT08.S3.NOx.~ NMHC.GT., na.action = na.exclude)
residuals(model_NMHC.GT.) %>% hist()
```
```{r}
residuals(model_NMHC.GT.) %>% boxplot()
```


```{r}
plot(model_NMHC.GT., which = c(1,2))
```

```{r message=FALSE, warning=FALSE}
summary(model_NMHC.GT.)
```

Prediction:

```{r paged.print=FALSE}
airq_data_2 <- airq_data %>% drop_na()
test_subset_NMHC.GT. <-  airq_data_2[which(row.names(airq_data_2) %in% sample(row.names(airq_data_2), 25, replace = FALSE)), c(3,7)]
test_NMHC.GT. <- data.frame(NMHC.GT. = test_subset_NMHC.GT.$NMHC.GT.)
test_subset_NMHC.GT.$pred_PT08.S3.NOx. <- predict(model_NMHC.GT., newdata = test_NMHC.GT.)
colnames(test_subset_NMHC.GT.) <- c('real_NMHC.GT.', 'real_PT08.S3.NOx.', 'pred_PT08.S3.NOx.')
head(test_subset_NMHC.GT.)
```
```{r message=FALSE, warning=FALSE}
R <- round(summary(model_NMHC.GT.)$adj.r.squared, digits = 3)
p <- round(summary(model_NMHC.GT.)$coefficients[2,4], digits = 3)
titl <- paste('R^2 =', as.character(R),', p-val =', as.character(p))
ggplot() + 
  geom_point(data = airq_data, aes(NMHC.GT., PT08.S3.NOx.)) + 
  geom_smooth(data = airq_data, aes(NMHC.GT., PT08.S3.NOx.), method = 'lm') +
  geom_point(data = test_subset_NMHC.GT., aes(real_NMHC.GT., real_PT08.S3.NOx.), color = 'red') + 
  geom_point(data = test_subset_NMHC.GT., aes(real_NMHC.GT., pred_PT08.S3.NOx.), color = 'green') +
  labs(title = titl)
```