---
output:
  pdf_document: default
  html_document: default
---
```{r}
library('ggplot2')
library('ggpubr')
```

```{r}
our_sample <- c(175, 176, 182, 165, 167, 172, 175,196, 158, 172)
```

\section{Measures of center}
\section{Mode, Median, Mean functions}

```{r}
mean(our_sample)
my_mean <- function(x){
  print(sum(x)/length(x))
}
my_mean(our_sample)
```

\textbf{mode with trimming}

```{r}
mean(our_sample, trim = 0.1)
```

```{r}
median(our_sample)
my_median <- function(x){
if (length(x) %% 2 == 0) {
  print((sort(x)[length(x)/2] + sort(x)[(length(x)/2) + 1])/2)
}
  else {
    print((sort(x)[(length(x)/2) + 1]))
    }
}
my_median(our_sample)
```

```{r}
my_mode <- function(x) {
t <- table(x)
print(as.numeric(names(t[t == max(t)])))
}
my_mode(our_sample)
```

\section{Histogram}

```{r}
ggplot() + 
  aes(our_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=mean(our_sample), color="red") +
  geom_vline(xintercept=median(our_sample), color="blue") +
  geom_vline(xintercept=my_mode(our_sample), color="green")
```

\section{Sample with the outlier}

```{r}
our_spoil_sample <- c(175, 176, 182, 165, 167, 172, 175,196, 158, 172, 235)
```

```{r}
mean(our_spoil_sample)
my_mean(our_spoil_sample)
```
```{r}
median(our_spoil_sample)
my_median(our_spoil_sample)
```

```{r}
my_mode(our_spoil_sample)
```

```{r}
ggplot() + 
  aes(our_spoil_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=mean(our_spoil_sample), color="red") +
  geom_vline(xintercept=median(our_spoil_sample), color="blue") +
  geom_vline(xintercept=my_mode(our_spoil_sample), color="green") 

```
\textbf{as median is the same as one of the mode value - there are only three lines}

\section{Measures of spread}
\section{Variance and Sd functions}
\textbf{R uses variance for unbiased estimators, I made the same}
```{r}
var(our_sample)
my_var <- function(x){
 n <- sapply(x, function(a)(a - mean(x))^2)
 print(sum(n)/(length(x)-1))
}
my_var(our_sample)
```
```{r}
sd(our_sample)
my_sd <- function(x){
 n <- sapply(x, function(a)(a - mean(x))^2)
 print((sum(n)/(length(x)-1))^0.5)
}
my_sd(our_sample)
```

\section{Histogram}

```{r}
ggplot() + 
  aes(our_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=c(median(our_sample) + 
                            IQR(our_sample),median(our_sample) - IQR(our_sample)), color="blue") +
  geom_vline(xintercept=c(mean(our_sample) + 
                            sd(our_sample),mean(our_sample) - sd(our_sample)), color="red") +
  geom_vline(xintercept=range(our_sample), color="green")

```

\section{Sample with the outlier}

```{r}
var(our_spoil_sample)
my_var(our_spoil_sample)
```


```{r}
sd(our_spoil_sample)
my_sd(our_spoil_sample)
```

```{r}
ggplot() + 
  aes(our_spoil_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=c(median(our_spoil_sample) + 
                            IQR(our_spoil_sample),median(our_spoil_sample) - 
                            IQR(our_spoil_sample)), color="blue") +
  geom_vline(xintercept=c(mean(our_spoil_sample) + 
                            sd(our_spoil_sample),mean(our_spoil_sample) - 
                            sd(our_spoil_sample)), color="red") +
  geom_vline(xintercept=range(our_spoil_sample), color="green")

```


\section{Check the properties for mean and sd for your sample}

```{r include=FALSE}
sum_table <- matrix(c(mean(our_sample), mean(our_sample - 100), mean(our_sample / 100), 
                      var(our_sample), var(our_sample - 100),var(our_sample / 100), 
                      sd(our_sample), sd(our_sample - 100), sd(our_sample / 100)), ncol = 3, byrow = TRUE)

colnames(sum_table) <- c("X","X-100","X/100")
rownames(sum_table) <- c("mean","var","sd")
sum_table <- as.table(sum_table)
```
```{r}
sum_table
```

```{r}
abs(sum(our_sample - mean(our_sample)) - 0) < 0.000000001
```
```{r}
a <- ggplot() + 
  aes(our_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=mean(our_sample-100), color="red") +
  ggtitle(label = 'Mean(x-100)')

b <- ggplot() + 
  aes(our_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=(mean(our_sample) -100), color="red") + 
  ggtitle(label = 'Mean(x)-100')

c <- ggplot() + 
  aes(our_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=(mean(our_sample/100)), color="red") + 
  ggtitle(label = 'Mean(x / 100)')
d <- ggplot() + 
  aes(our_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=(mean(our_sample)/100), color="red") + 
  ggtitle(label = 'Mean(x) / 100)')

ggarrange(a, b, c, d, ncol = 2, nrow = 2)
```

```{r}
e <- ggplot() + 
  aes(our_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=(var(our_sample - 100)), color="red") + 
  ggtitle(label = 'Var(x - 100)')
f <- ggplot() + 
  aes(our_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=(var(our_sample)), color="red") + 
  ggtitle(label = 'Var(x)')
g <- ggplot() + 
  aes(our_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=(var(our_sample/100)), color="red") + 
  ggtitle(label = 'Var(x / 100)')
k <- ggplot() + 
  aes(our_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=(var(our_sample/10000)), color="red") + 
  ggtitle(label = 'Var(x) / 10000')
ggarrange(e, f, g, k, ncol = 2, nrow = 2)
```

```{r include=FALSE}
l <- ggplot() + 
  aes(our_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=(sd(our_sample / 100)), color="red") + 
  ggtitle(label = 'Sd(x / 100)')
m <- ggplot() + 
  aes(our_sample) + 
  geom_histogram(binwidth=2, colour="black", fill="white") +
  geom_vline(xintercept=(sd(our_sample)/100), color="red") + 
  ggtitle(label = 'Sd(x) / 100')


```
```{r}
ggarrange(l,m, ncol = 2, nrow = 1)
```

\section{Normal distribution}
\textbf{For the population N(175, 10) find the probability to be less than 156cm}
```{r}
pnorm(156, 175, 10)
```

\textbf{For the population N(175, 10) find the probability to be more than 198 cm}
```{r}
pnorm(198, 175, 10, lower.tail = FALSE)
```

\textbf{For the population N(175, 10) find the probability to be between 168 and 172 cm}
```{r}
pnorm(172, 175, 10) - pnorm(168, 175, 10)
```

\section{Standard normal distribution}
```{r}
pnorm(1) - pnorm(-1) # 68% of the data is within 1 standard deviation 
pnorm(2) - pnorm(-2) # 95% of the data is within 2 standard deviations
pnorm(3) - pnorm(-3) # 99.7% of the data is within 3 standard deviations
```
\section{Generate sample using rnorm() from N(175, 10), find mean and sd}

```{r}
set.seed(42)
norm_sample <- rnorm(1000,175,10)
mean(norm_sample)
sd(norm_sample)
```

\section{Standardize}

```{r}
stand_norm_sample <- scale(norm_sample)
mean(norm_sample)
sd(norm_sample)
```

\section{Central Limit Theorem}
 
```{r echo=TRUE}
set.seed(42)
large_population <- rnorm(1000000)
my_samples_ten <- replicate(10, sample(large_population, 30)) #k  = 10
means_ten <- colMeans(my_samples_ten)
my_samples_fifty <- replicate(50, sample(large_population, 30)) #k = 50
means_fifty <- colMeans(my_samples_fifty)
my_samples_oneh <- replicate(100, sample(large_population, 30)) #k = 100
means_oneh <- colMeans(my_samples_oneh)
my_samples_fifh <- replicate(500, sample(large_population, 30)) #k = 500
means_fifh <- colMeans(my_samples_fifh)

se <- function(x) sqrt(var(x)/length(x))
means_table <- matrix(c(mean(means_ten), sd(means_ten), se(means_ten), 
                      mean(means_fifty), sd(means_fifty), se(means_fifty), 
                      mean(means_oneh), sd(means_oneh), se(means_oneh),
                      mean(means_fifh), sd(means_fifh), se(means_fifh)), ncol = 3, byrow = TRUE)

colnames(means_table) <- c("mean","sd","SE")
rownames(means_table) <- c("10","50","100", "500")
means_table <- as.table(means_table)
```
```{r}
means_table
```


```{r}
n <- ggplot() + 
  aes(means_ten) + 
  geom_histogram(binwidth=0.125, colour="black", fill="white") +
  geom_vline(xintercept=mean(means_ten), color="red") +
  geom_vline(xintercept=c(mean(means_ten) + sd(means_ten), 
                          mean(means_ten) - sd(means_ten)), color="blue") +
  geom_vline(xintercept=c(mean(means_ten) + se(means_ten), 
                          mean(means_ten) - se(means_ten)), color="green") +
  ggtitle(label = 'k = 10')

o <- ggplot() + 
  aes(means_fifty) + 
  geom_histogram(binwidth=0.125, colour="black", fill="white") +
  geom_vline(xintercept=mean(means_fifty), color="red") +
  geom_vline(xintercept=c(mean(means_fifty) + sd(means_fifty), 
                          mean(means_fifty) - sd(means_fifty)), color="blue") +
  geom_vline(xintercept=c(mean(means_fifty) + se(means_fifty),
                          mean(means_fifty) - se(means_fifty)), color="green") +
  ggtitle(label = 'k = 50')

p <- ggplot() + 
  aes(means_oneh) + 
  geom_histogram(binwidth=0.125, colour="black", fill="white") +
  geom_vline(xintercept=mean(means_oneh), color="red") +
  geom_vline(xintercept=c(mean(means_oneh) + sd(means_oneh), 
                          mean(means_oneh) - sd(means_oneh)), color="blue") +
  geom_vline(xintercept=c(mean(means_oneh) + se(means_oneh),
                          mean(means_oneh) - se(means_oneh)), color="green") +
  ggtitle(label = 'k = 100')

k <- ggplot() + 
  aes(means_fifh) + 
  geom_histogram(binwidth=0.125, colour="black", fill="white") +
  geom_vline(xintercept=mean(means_fifh), color="red") +
  geom_vline(xintercept=c(mean(means_fifh) + sd(means_fifh), 
                          mean(means_fifh) - sd(means_fifh)), color="blue") +
  geom_vline(xintercept=c(mean(means_fifh) + se(means_fifh),
                          mean(means_fifh) - se(means_fifh)), color="green") +
  ggtitle(label = 'k = 500')

ggarrange(n, o, p, k, ncol = 2, nrow = 2)
```

