---
title: "Task 2, multiple regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(corrplot)
library(car)
library(plotly)
```

```{r}
airq_data <- read.table('/home/marina/Загрузки/AirQualityUCI.csv', sep = ';', header = TRUE, dec = ",")
```

```{r}
airq_data <- airq_data %>% select_if(~sum(!is.na(.)) > 0)  %>% drop_na()
airq_data[c('CO.GT.', 'C6H6.GT.', 'T', 'RH', 'AH')] <-  sapply(airq_data[c('CO.GT.', 'C6H6.GT.', 'T', 'RH', 'AH')], as.numeric)
airq_data <- airq_data[, c(3:15)]
```

```{r}
airq_data <- sapply(airq_data, function(x){ifelse (x == -200, NA, x)})
airq_data <- as.data.frame(airq_data)
```

```{r}
cor_data <- cor(airq_data, method = 'spearman', use = 'pairwise.complete.obs')
corrplot(cor_data, method = 'circle')
```

•  Exploring different models

1) I decided to take the best predictors for linear model and combine them into one multiple

```{r}
first_model <- lm(PT08.S3.NOx. ~ PT08.S2.NMHC. + PT08.S1.CO. + PT08.S5.O3., data = airq_data, na.action = na.exclude)
summary(first_model)
```

```{r}
vif(first_model)
```

Looking good! Let's check the residuals:

```{r}
plot(first_model, which = c(1,2))
```

```{r}
residuals(first_model) %>% hist()
```

2) Ok, let's check if we also use the interaction between the variables:

```{r}
second_model <- lm(PT08.S3.NOx. ~ PT08.S2.NMHC. * PT08.S1.CO. * PT08.S5.O3., data = airq_data, na.action = na.exclude)
summary(second_model)
```

```{r}
vif(second_model)
```

R^2 is bigger, but now there is a multicollinearity

3) Also, I want to check if we can make a model with 2 preditors. Maybe it would be better.
This model was the best one:

```{r}
third_model <- lm(PT08.S3.NOx. ~ PT08.S2.NMHC. + PT08.S5.O3., data = airq_data, na.action = na.exclude)
summary(third_model)
```

```{r}
vif(third_model)
```

```{r}
plot(third_model, which = c(1,2))
```

```{r}
residuals(third_model) %>% hist()
```

Looks good, R^2 is the same as with 3 predictors.

4) Also, we can try to make a model with 'step' function. It is not working with NA's, so to use it I should clean a data.
I delete the NMHC.GT. column to save more data. So, in the new data without NA I have 6841 rows.

```{r include=FALSE}
airq_data_2 = airq_data[, -3] %>% drop_na()
full_model <- lm(PT08.S3.NOx. ~ (.)^2, data = airq_data_2)
null_model <- lm(PT08.S3.NOx. ~ 1, data = airq_data_2)
fourth_model <- step(null_model, scope = list(lower = null_model, upper = full_model), direction = "forward")
```

```{r}
summary(fourth_model)
```

```{r}
vif(fourth_model)
```

```{r}
plot(fourth_model, which = c(1,2))
```

```{r}
residuals(fourth_model) %>% hist()
```

R^2 is bigger, but there is a lot of multicollinearity and the residuals looking not very good.

•  So, I decided to choose the third model (with 2 predictors) as the best one. Now, I want to check if the transformation will improve the model.

```{r}
R <- NULL
for (i in -5:5){
  one <- airq_data$PT08.S2.NMHC.
  two <-  airq_data$PT08.S5.O3.
  if (i < 0){
    one <- one^(i) *(-1)
    two <- two^(i) *(-1)
  }
  else if (i == 0) {
    one <- log(one)
    two <-  log(two)
  }
  else {
    one <- one^i 
    two <-  two^i 
  }
  model <- lm(PT08.S3.NOx. ~ one + two, data = airq_data, na.action = na.exclude)
  R <- c(R, summary(model)$adj.r.squared)
}
transform <- data.frame('p' = c(-5:5), 'R^2' = R)
```

```{r}
plot(transform, type = 'b')
```

```{r}
transform[max(transform),]
```

•  Making the best model and predicting the values

```{r}
airq_data$PT08.S2.NMHC. <- (airq_data$PT08.S2.NMHC.)^-1 * (-1)
airq_data$PT08.S5.O3. <- (airq_data$PT08.S5.O3.)^-1 * (-1)
final_model <- lm(PT08.S3.NOx. ~ PT08.S2.NMHC. + PT08.S5.O3., data = airq_data, na.action = na.exclude)
```

```{r}
summary(final_model)
```

```{r}
vif(final_model)
```

```{r}
plot(final_model, which = c(1,2))
```

```{r}
residuals(final_model) %>% hist()
```

•  Prediction:

```{r}
test_subset_multiple <-  airq_data[which(row.names(airq_data) %in% sample(row.names(airq_data), 50, replace = FALSE)), c(5,10,7)]
test_multiple <- data.frame(PT08.S2.NMHC. = test_subset_multiple$PT08.S2.NMHC., PT08.S5.O3. = test_subset_multiple$PT08.S5.O3.)
test_subset_multiple$pred_PT08.S3.NOx. <- predict(final_model, newdata = test_multiple)
colnames(test_subset_multiple) <- c('real_PT08.S2.NMHC.','real_PT08.S5.O3.', 'real_PT08.S3.NOx.', 'pred_PT08.S3.NOx.')
```

```{r}
df <- data.frame('y' = airq_data$PT08.S3.NOx. , 'x1' = airq_data$PT08.S2.NMHC., 'x2' = airq_data$PT08.S5.O3.)

### Estimation of the regression plane
mod <- lm(y ~ x1+x2, data = df)
cf.mod <- coef(mod)
df <- df %>% drop_na()
### Calculate z on a grid of x-y values
PT08.S2.NMHC. <- seq(min(df$x1),max(df$x1),length.out=25)
PT08.S5.O3. <- seq(min(df$x2),max(df$x2),length.out=25)
z <- t(outer(PT08.S2.NMHC., PT08.S5.O3., function(x,y) cf.mod[1]+cf.mod[2]*x+cf.mod[3]*y))
#### Draw the plane with "plot_ly" and add points with "add_trace"

R <- round(summary(final_model)$adj.r.squared, digits = 3)
p <- round(summary(final_model)$coefficients[2,4], digits = 3)
titl <- paste('R^2 =', as.character(R),', p-val =', as.character(p))
 

plot_ly(x=~PT08.S2.NMHC., y=~PT08.S5.O3., z=~z, type="surface", opacity=0.7)  %>%
  add_trace(data = test_subset_multiple, x = ~ real_PT08.S2.NMHC., y = ~ real_PT08.S5.O3., 
            z = ~  pred_PT08.S3.NOx., type="scatter3d", size = 1, name= "Predicted", opacity=0.8) %>%
  add_trace(data = test_subset_multiple, x = ~ real_PT08.S2.NMHC., y = ~ real_PT08.S5.O3., 
            z = ~  real_PT08.S3.NOx., type="scatter3d", size = 1, name="Real", opacity=0.8) %>% 
  layout(title = titl)
```

