---
title: "HW_1 Grigoreva Elizaveta"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(ggpubr)
```

#1. Measures of center
# 1.0 create own sample or use given vector and write mode, median, mean functions/one-liners
```{r}
x <- c(175, 176, 180, 165, 167, 172, 175, 146, 158, 178)

#Get mode
mymode <- function(x){
  un <- unique(x)
  r <- tabulate(match(x, un))
  return(un[r == max(r)])
}

#Median
mymedian <- function(x) {
  n <- length(x)
  s <- sort(x)
  ifelse(n%%2==1,s[(n+1)/2],mean(s[n/2+0:1]))
}
#Mean
mymean <- sum(x)/length(x)
```
# 1.1 calculate mode, median and mean for the sample.
#Compare results for own and  built-ins for median and mean
```{r}
mymode(x)
mymedian(x)
median(x)
mymean
mean(x)
```
# 1.2 visualize histogram with 3 vertical lines for measures of center
```{r}
ggplot(as.data.frame(x), aes(x = x)) +
  geom_histogram() +
  geom_vline(xintercept = mymean, color = 'black') +
  geom_vline(xintercept = mymedian(x), color = 'red') +
  geom_vline(xintercept = mymode(x), color = 'green')
```



# # 1.3 spoil your sample with the outlier - repeat steps 1.1 and 1.2
```{r}
x[length(x) + 1] <- 15
## 1.1. repeat
mymode(x)
mymedian(x)
median(x)
mymean2 <- sum(x)/length(x)
mymean2
mean(x)
```

# Vizualize it
```{r}
ggplot(as.data.frame(x), aes(x = x)) +
  geom_histogram() +
  geom_vline(xintercept = mymean2, color = 'black') +
  geom_vline(xintercept = mymedian(x), color = 'red') +
  geom_vline(xintercept = mymode(x), color = 'green')
```


#Measures of spread
# 2.0 write the functions/one-liners for variance and sd, calculate result, compare with the built-ins
```{r}
x <- c(175, 176, 182, 165, 167, 172, 175, 196, 158, 172)
var_one_line <- mean((x-mean(x))^2)
sd_one_line <- sqrt(sum((x-mean(x))^2/(length(x)-1)))
var(x)
var_one_line
sd(x)
sd_one_line
```
## 2.1 visualize with the box plot.
# Add horizontal lines for range, IQR, 1-sd borders (use built-ins)
```{r}
ggplot(as.data.frame(x), aes(y = x)) +
  geom_boxplot() +
  geom_hline(yintercept = min(x), color = 'red') +
  geom_hline(yintercept = max(x), color = 'red') +
  geom_hline(yintercept = quantile(x, 3/4), color = 'green') +
  geom_hline(yintercept = quantile(x, 1/4), color = 'green')

```
# 2.2 spoil your sample with the outlier, repeat step 2.1 and back vector
```{r}
x[length(x) + 1] <- 15
ggplot(as.data.frame(x), aes(y = x)) +
  geom_boxplot() +
  geom_hline(yintercept = min(x), color = 'red') +
  geom_hline(yintercept = max(x), color = 'red') +
  geom_hline(yintercept = quantile(x, 3/4), color = 'green') +
  geom_hline(yintercept = quantile(x, 1/4), color = 'green')

x <- x[1:length(x) - 1]
```
# 3. Properties
# 3.0 check the properties for mean and sd for your sample
```{r}
x <- c(175, 176, 182, 165, 167, 172, 175,196, 158, 172)
var(x - 100)
var(x)
var(x / 100)
var(x) / 10000
sd(x / 100)
sd(x) / 100
abs(sum(x) -mean(x) -0 ) < 0.000000001
```
# 3.1 visualize result tabularly and graphically (maybe with facetting free scales?)
```{r}
#Vizualize tabulary
properties_table <- matrix(c(mean(x), mean(x - 100), mean(x/ 100), 
                      var(x), var(x - 100),var(x / 100), 
                      sd(x), sd(x - 100), sd(x / 100)), ncol = 3, byrow = TRUE)

colnames(properties_table ) <- c("x","x-100","x/100")
rownames(properties_table ) <- c("mean","var","sd")
as.table(properties_table)
#Vizualize for mean
a <- ggplot() + 
  aes(x) + 
  geom_histogram(colour="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 300, by = 8)) +
  geom_vline(xintercept=mean(x-100), color="red") +
  ggtitle(label = 'Mean(x-100)')
b <- ggplot() + 
  aes(x) + 
  geom_histogram( colour="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 300, by = 8)) +
  geom_vline(xintercept=(mean(x/100)), color="red") + 
  ggtitle(label = 'Mean(x / 100)')
c <- ggplot() + 
  aes(x) + 
  geom_histogram(colour="black", fill="white") +
  geom_vline(xintercept=(mean(x)), color="red") + 
  scale_x_continuous(breaks = seq(0, 300, by = 8)) +
  ggtitle(label = 'Mean(x)')
ggarrange(a, b, c,  ncol = 3, nrow = 1)

#Vizualize for var
d <- ggplot() + 
  aes(x) + 
  geom_histogram(colour="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 300, by = 8)) +
  geom_vline(xintercept=var(x-100), color="red") +
  ggtitle(label = 'sd(x-100)')
e <- ggplot() + 
  aes(x) + 
  geom_histogram( colour="black", fill="white") +
  geom_vline(xintercept=(var(x/100)), color="red") +
  scale_x_continuous(breaks = seq(0, 300, by = 8)) +
  ggtitle(label = 'sd(x / 100)')
f <- ggplot() + 
  aes(x) + 
  geom_histogram( colour="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 300, by = 8)) +
  geom_vline(xintercept=(var(x)), color="red") + 
  ggtitle(label = 'sd(x)')
ggarrange(d, e, f,  ncol = 3, nrow = 1)

#Vizualize for sd
g <- ggplot() + 
  aes(x) + 
  geom_histogram(colour="black", fill="white") +
  geom_vline(xintercept=sd(x-100), color="red") +
  scale_x_continuous(breaks = seq(0, 300, by = 8)) +
  ggtitle(label = 'sd(x-100)')
h <- ggplot() + 
  aes(x) + 
  geom_histogram( colour="black", fill="white") +
  geom_vline(xintercept=(sd(x/100)), color="red") + 
  scale_x_continuous(breaks = seq(0, 300, by = 8)) +
  ggtitle(label = 'sd(x / 100)')
i <- ggplot() + 
  aes(x) + 
  geom_histogram(colour="black", fill="white") +
  geom_vline(xintercept=(sd(x)), color="red") + 
  scale_x_continuous(breaks = seq(0, 300, by = 8)) +
  ggtitle(label = 'sd(x)')
ggarrange(g, h, i,  ncol = 3, nrow = 1)

```
# 4 Normal Distribution
# 4.0 for the population N(175, 10) find the probability to be:
# less than 156cm,
# more than 198,
# between 168 and 172 cm
```{r}
pnorm(156,175,10, lower.tail = TRUE)
pnorm(198,175,10, lower.tail = FALSE)
pnorm(168,175,10, lower.tail = FALSE)-pnorm(172, mean = 175, sd = 10, lower.tail = TRUE)

```
## Standard normal distribution
# 4.1 check the properties of 1-2-3-sd's for standard normal distribution using pnorm()
```{r}
pnorm(1) - pnorm(-1) # 68,2% 
pnorm(2) - pnorm(-2) # 95,4% 
pnorm(3) - pnorm(-3)# 99.7%
```
# 4.3 standardize, find the same
```{r}
set.seed(42)
x <- rnorm(1000,175,10)
mean(x)
sd(x)
x1 <- (x-mean(x))/sd(x)
mean(x1)
sd(x1)
x <- rnorm(1000, mean = 0, sd = 1)
mean(x)
sd(x)
```

# 5. Central Limit Theorem
# 5.0 Generate large population (n ~ 100 000 - 1 000 000) distributed as N(0, 1) 
# Sample from population k observations for 30 times - you will have set of 30 samples.
# For each sample calculate mean. For the set calculate means of means, sd of means, SE.
# Create table with k, mean of means, sd of means, SE.
# Visualize distribution of means with histogram and lines for mean of means and SE.
# 5.1 k = 10
# 5.2 k = 50
# 5.3 k = 100
# 5.4 k = 500
# 5.5 Compare results
```{r}
set.seed(42)
x <- rnorm(1000000, mean = 0, sd = 1)
#10,50,100,500
k_10 <- replicate(30, sample(x, 10))
k_50 <- replicate(30,sample(x,50))
k_100 <- replicate(30,sample(x,100))
k_500 <- replicate(30,sample(x,500))
#mean
means <- function(k){
  m <- c()
  for (i in 1:ncol(k)) {
    m[i] <- mean(k[,i])
  }
  return(m)
}
SE <- function(k){
  return(sd(k)/sqrt(length(k)))
}
means_of_mean_K_10 <- means(k_10)
mean(means_of_mean)
sd(means_of_mean_K_10)
sd(means_of_mean)
SE(means_of_mean_K_10)
#50
means_of_mean_K_50 <- means(k_50)
mean(means_of_mean_K_50)
sd(means_of_mean_K_50)
 SE(means_of_mean_K_50)
#100 
means_of_mean_K_100 <- means(k_100)
mean(means_of_mean_K_100)
sd(means_of_mean_K_100)
 SE(means_of_mean_K_100)
#500
means_of_mean_K_500 <- means(k_500)
mean(means_of_mean_K_500)
sd(means_of_mean_K_500)
SE(means_of_mean_K_500)

#Create table
table_samples <- matrix(c(mean(means_of_mean_K_10),sd(means_of_mean_K_10),SE(means_of_mean_K_10),
                    mean(means_of_mean_K_50), sd(means_of_mean_K_50), SE(means_of_mean_K_50),
                    mean(means_of_mean_K_100), sd(means_of_mean_K_100), SE(means_of_mean_K_100),
                    mean(means_of_mean_K_500), sd(means_of_mean_K_500), SE(means_of_mean_K_500)),ncol=3)

colnames(table_samples) <- c("mean","sd","SE")
rownames(table_samples) <- c("10","50","100", "500")
table_samples <- as.table(table_samples)
table_samples 
#Vizualizing distribution
a <- ggplot() + 
aes(x) + 
  geom_histogram(colour="black", fill="white") +
  geom_vline(xintercept=mean(means_of_mean_K_10), color="red") +
  geom_vline(xintercept=sd(means_of_mean_K_10),color="blue")+
  geom_vline(xintercept=SE(means_of_mean_K_10), color='green')+
  ggtitle(label = 'mean,sd,SE,k=10')
 
b <- ggplot() + 
aes(x) + 
  geom_histogram(colour="black", fill="white") +
  geom_vline(xintercept=mean(means_of_mean_K_50), color="red") +
  geom_vline(xintercept=sd(means_of_mean_K_50),color="blue")+
  geom_vline(xintercept=SE(means_of_mean_K_50), color='green')+
  ggtitle(label = 'mean,sd,SE,k=10')
  
c <- ggplot() + 
aes(x) + 
  geom_histogram(colour="black", fill="white") +
  geom_vline(xintercept=mean(means_of_mean_K_100), color="red") +
  geom_vline(xintercept=sd(means_of_mean_K_100),color="blue")+
  geom_vline(xintercept=SE(means_of_mean_K_100), color='green')+
  ggtitle(label = 'mean,sd,SE,k=10')

e <- ggplot() + 
aes(x) + 
  geom_histogram(colour="black", fill="white") +
  geom_vline(xintercept=mean(means_of_mean_K_500), color="red") +
  geom_vline(xintercept=sd(means_of_mean_K_500),color="blue")+
  geom_vline(xintercept=SE(means_of_mean_K_500), color='green')+
  ggtitle(label = 'mean,sd,SE,k=10')
ggarrange(a, b, c,d,e,  ncol = 2, nrow = 3)
```












