---
title: "HW01"
author: "Maria Firuleva"
date: '10/10/2019 г '
output: html_document
---


## Load libraries

```{r libraries, message=F}
library(dplyr)
library(tidyverse)
```

## Function

There are several arguments for *dataSubset* function:

  * **data**: input dataframe, required argument;
  
  * **colNames**: vector of column names which you want to save in result dataframe. It is first step of dataframe filtration and, by default, it includes all column names in input dataframe;
  
  * **colIndexes**: vector of column indexes which you want to save in result dataframe *after* filtering using colNames; make sure there is no conflict between these two steps; by default, it includes all column indexes in input dataframe;
  
  * **colLogicVector**: logical vector for column subsetting which you want to save in result dataframe; this parameter will be used after filtration using **colNames** and **colIndexes**; by default, it parameter is *NULL* and, after columns filtration using **colNames** and **colIndexes**, it will be redefined as vector with only *TRUE* values;
  
  * **rowNames**: vector of row names which you want to save in result dataframe; by default, it includes all row names in input dataframe;
  
  * **rowIndexes**: vector of row indexes which you want to save in result dataframe; by default, it includes all row indexes in input dataframe.
  

Function output is a named list with several elements:

  * **Data**: filtered input dataframe;
  
  * **columnSums**: named numeric vector contains column sums for numeric column in filtered dataframe, NA will be removed;
  
  * **columnSDs**: named numeric vector contains column SDs for numeric columnss in filtered dataframe, NA will be removed;
  
  * **columnMeans**: named numeric vector contains column means for numeric columns in filtered dataframe, NA will be removed;
  
  * **freqTable**: frequency table for all non-numeric columns in filtered dataframe.


```{r function}
dataSubset <- function(data, colNames = colnames(data), colIndexes = c(1:ncol(data)), colLogicVector = NULL, rowNames = rownames(data), rowIndexes = c(1:nrow(data))) {
  data <- select(data, colNames)
  data <- select(data, colIndexes)
  if (is.null(colLogicVector)) {
    colLogicVector <- c(rep(T, ncol(data)))
  }
  data <- select(data, which(colLogicVector))
  data <- data %>% rownames_to_column("target") %>% filter(rownames(data) %in% rowNames) %>% remove_rownames %>% column_to_rownames(var="target")
  data <- data %>% rownames_to_column("target") %>% slice(rowIndexes) %>% remove_rownames %>% column_to_rownames(var="target")
  columnSums <- list()
  columnMeans <- list()
  columnSDs <- list()
  freqTable <- list()
  for (col in seq_along(1:ncol(data))) {
    if (is.numeric(data[,col])) {
      columnSDs[[colnames(data)[col]]] <- sd(data[,col], na.rm = T)
      columnMeans[[colnames(data)[col]]] <- mean(data[,col], na.rm = T)
      columnSums[[colnames(data)[col]]] <- sum(data[,col], na.rm = T)
    } else{
      freqTable[[colnames(data)[col]]] <- c(table(data[,col]))
    }
  }
  data <- list(data, columnSums, columnSDs, columnMeans, freqTable)
  names(data) <- c("Data", "columnSums", "columnSDs", "columnMeans", "freqTable")
  data
}
```

Let's look on some examples.

## mtcars

```{r mtcars}
result <- dataSubset(mtcars, colNames = c("disp", "hp", "cyl"),  colIndexes = c(1:3), rowNames = c("Mazda RX4", "Merc 450SE", "Fiat 128", "Ford Pantera L"), rowIndexes = c(1:2))
str(result$Data)
result$columnSums
result$columnSDs
result$columnMeans
result$freqTable
```

In this case, there is an error because we want to save column 4, but our prefiltered dataset contains only 3 columns: disp, hp and cyl.

```{r mtcars_error, error=TRUE}
result <- dataSubset(mtcars, colNames = c("disp", "hp", "cyl"),  colIndexes = c(1:4), rowNames = c("Mazda RX4", "Merc 450SE", "Fiat 128", "Ford Pantera L"), rowIndexes = c(1:2))
```



## iris

```{r iris}
result <- dataSubset(iris, colIndexes = c(1:3, 5), rowIndexes = c(1:2, 89:95, 110:130))
str(result$Data)
result$columnSums
result$columnSDs
result$columnMeans
result$freqTable
```

What about NA in factor variable?

```{r iris_with_na}
iris$Species[1] <- NA
result <- dataSubset(iris, colIndexes = c(1:3, 5), rowIndexes = c(1:2, 89:95, 110:130))
result$freqTable
```


## airquality

```{r airquality}
result <- dataSubset(airquality, rowIndexes = c(1:25), colLogicVector = c(T, F, T, F))
str(result$Data)
result$columnSums
result$columnSDs
result$columnMeans
result$freqTable
```

