---
title: "HW02"
author: "Maria Firuleva"
date: '10/10/2019 г '
output: html_document
---


## Function

There are several arguments for *dataSubset* function:

  * **data**: input dataframe, required argument;
  
  * **colNames**: vector of column names which you want to save in result dataframe. It is first step of dataframe filtration and, by default, it includes all column names in input dataframe;
  
  * **colIndexes**: vector of column indexes which you want to save in result dataframe *after* filtering using colNames; make sure there is no conflict between these two steps; by default, it includes all column indexes in input dataframe;
  
  * **colLogicVector**: logical vector for column subsetting which you want to save in result dataframe; this parameter will be used after filtration using **colNames** and **colIndexes**; by default, it parameter is *NULL* and, after columns filtration using **colNames** and **colIndexes**, it will be redefined as vector with only *TRUE* values;
  
  * **rowNames**: vector of row names which you want to save in result dataframe; by default, it includes all row names in input dataframe;
  
  * **rowIndexes**: vector of row indexes which you want to save in result dataframe; by default, it includes all row indexes in input dataframe;
  
  * **calculations**: vector with combinations of "columnSums", "columnMeans" and "columnSDs" strings or NULL.
  

Function output is a named list with several elements:

  * **Data**: filtered input dataframe;
  
  * **columnSums**: named numeric vector contains column sums for numeric column in filtered dataframe, NA will be removed (optional, depends on *calculations* argument);
  
  * **columnSDs**: named numeric vector contains column SDs for numeric columnss in filtered dataframe, NA will be removed (optional, depends on *calculations* argument);
  
  * **columnMeans**: named numeric vector contains column means for numeric columns in filtered dataframe, NA will be removed (optional, depends on *calculations* argument);
  
  * **freqTable**: frequency table for all non-numeric columns in filtered dataframe.


```{r function}
library(dplyr)
dataSubset <- function(data, colNames = colnames(data), colIndexes = c(1:ncol(data)), colLogicVector = NULL, rowNames = rownames(data), rowIndexes = c(1:nrow(data)), calculations = c("columnSums", "columnMeans", "columnSDs"), splitted_factor = NULL) {
  if (!is.null(calculations) & sum(!calculations %in% c("columnSums", "columnMeans", "columnSDs"))) {
    stop('calculations argument can be a vector with combinations of columnSums, columnMeans and columnSDs strings or NULL. Perhaps you put something unacceptable string in this vector.')
  }
  data <- data[colNames]
  if (max(colIndexes) > ncol(data)) {
    stop('You must specify colIndexes after filtering using colNames; make sure there is no conflict between these arguments.')
  }
  data <- data[colIndexes]
  init_colnames <- colnames(data)
  if (is.null(colLogicVector)) {
    colLogicVector <- c(rep(T, ncol(data)))
  }
  if (length(colLogicVector) > ncol(data)) {
    stop('You must specify colLogicVector after filtering using colNames and colIndexes; make sure there is no conflict between these arguments.')
  }
  data <- data[which(colLogicVector)]
  init_colnames <- colnames(data)
  data <- data[rowNames,] %>% as.data.frame() %>% magrittr::set_colnames(init_colnames)
  if (max(rowIndexes) > nrow(data)) {
    stop('You must specify rowIndexes after filtering using rowNames; make sure there is no conflict between these arguments.')
  }
  data <- data[rowIndexes,] %>% as.data.frame() %>% magrittr::set_colnames(init_colnames)
  non_numeric_data <- data[!sapply(data, is.numeric)]
  result <- list()
  result$data <- data
  if (!is.null(splitted_factor)) {
    data <- split(data, data[splitted_factor])
    data <- lapply(data, function(x) x[sapply(x, is.numeric)])
    if ("columnSums" %in% calculations) {
      columnSums <- lapply(data, colSums, na.rm = T)
      result$columnSums <- columnSums
    }
    if ("columnMeans" %in% calculations) {
      columnMeans <- lapply(data, colMeans, na.rm = T)
      result$columnMeans <- columnMeans
    }
    if ("columnSDs" %in% calculations) {
      columnSDs <- lapply(data, function(x) lapply(x, sd, na.rm = T))
      result$columnSDs <- columnMeans
    }
  } else {
    data <- lapply(data, function(x) x[sapply(x, is.numeric)])
    if ("columnSums" %in% calculations) {
      columnSums <- lapply(data, sum, na.rm = T)
      result$columnSums <- columnSums
    }
    if ("columnMeans" %in% calculations) {
      columnMeans <- lapply(data, mean, na.rm = T)
      result$columnMeans <- columnMeans
    }
    if ("columnSDs" %in% calculations) {
      columnSDs <- lapply(data, function(x) lapply(x, sd, na.rm = T))
      result$columnSDs <- columnMeans
    }
  }
  freqTable <- apply(non_numeric_data, 2, table)
  result$freqTable <- freqTable
  result
}
```

Let's look on some examples.

## Nikita's issue

```{r fixed}
dataSubset(iris, colIndexes = 1, rowIndexes = c(1:12))
```


```{r iris}
result <- dataSubset(iris, colIndexes = c(1:3, 5), rowIndexes = c(1:2, 89:95, 110:130), splitted_factor = "Species")
str(result$data)
result$columnSums
result$columnSDs
result$columnMeans
result$freqTable
```

What about NA in factor variable?

```{r iris_with_na}
iris$Species[1] <- NA
result <- dataSubset(iris, colIndexes = c(1:3, 5), rowIndexes = c(1:2, 89:95, 110:130), splitted_factor = "Species", calculations="columnSums")
str(result)
result$freqTable
```


## airquality

```{r airquality}
result <- dataSubset(airquality, rowIndexes = c(1:25), colLogicVector = c(T, F, T, F))
str(result)
result$columnSums
result$columnSDs
result$columnMeans
result$freqTable
```

## mtcars

In this case, we don't perform any calculations, and final result objects is a combination of two lists: filtered data and freqTable (which empty, in our case).

```{r mtcars}
result <- dataSubset(mtcars, colNames = c("disp", "hp", "cyl"),  colIndexes = c(1:3), rowNames = c("Mazda RX4", "Merc 450SE", "Fiat 128", "Ford Pantera L"), rowIndexes = c(1:2), calculations = NULL)
str(result)
```

What if we use wrong values for calculations argument?

```{r mtcars_error, error=TRUE}
result <- dataSubset(mtcars, colNames = c("disp", "hp", "cyl"),  colIndexes = c(1:4), rowNames = c("Mazda RX4", "Merc 450SE", "Fiat 128", "Ford Pantera L"), rowIndexes = c(1:2), calculations = c("columnSums", "columnMeans", "columnSd"))
```