---
title: 'Weather: data clining'
author: "Maria Firuleva"
date: '7 November 2019 г '
output: html_document
---


## HW4: weather

# Load libraries and define functions

```{r loading, message=F}
library(data.table)
library(dplyr)
library(ggplot2)
library(tidyr)
library(Hmisc)
```


```{r functions}
getMode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}

getNApercentage <- function(data) {
  data <- as.data.frame(colMeans(is.na(data)) * 100)
  data$measure <- rownames(data)
  colnames(data)[1] <- 'NA_percentage'
  ggplot(data %>% filter(NA_percentage > 10), aes(x=measure, weight=NA_percentage)) + geom_bar(color="black", fill="white")+labs(x = "Measure", y = "NA percentage", title="Measurements with NA percentage >= 10%")+theme_classic()+ theme(aspect.ratio = 1)
}

removeOutliers <- function(x){
    q <- quantile(x, 0.25) + quantile(x, 0.75)
    return(x[abs(x - q/2) <= 2*IQR(x)])
}


replaceNA <- function(x, treshold=1){
  totalUniqValues <- length(unique(na.omit(x)))
  outNumber <- totalUniqValues - length(unique(removeOutliers(na.omit(x))))
  ifelse(outNumber >= treshold, x <- replace(x, is.na(x), mean(x, na.rm = TRUE)), x <- replace(x, is.na(x), median(x, na.rm = TRUE))) 
  x
}


```

# Load data, reshape it

```{r reshaping}
weather <- readRDS("~/master/R/Lessons/weather.rds")
weather <- select(weather, -X)
weather <- unite(weather, period, year, month)
weather <- melt(setDT(weather), 
         measure.vars = patterns("^X"), 
         variable.name = "Day")

weather$Day <- gsub('X', '', weather$Day)
weather <- spread(weather, measure, value)
weather <- as.data.frame(weather)
weather$period <- as.factor(weather$period)
weather$Events <- as.factor(weather$Events)
weather[,c(2:3, 5:ncol(weather))] <- apply(weather[,c(2:3,5:ncol(weather))], 2, as.numeric)
tbl_df(weather)
```


# Test whether data contains columns with NA >= 10% or not

```{r estimateNA}
getNApercentage(weather)
```

There are two numerical columns with NA percentages more than 10%.

# Visualize distributions of all numeric continous columns (without Day column)

```{r plots}
hist.data.frame(na.omit(select_if(weather, is.numeric))[2:5])
hist.data.frame(na.omit(select_if(weather, is.numeric))[6:10])
hist.data.frame(na.omit(select_if(weather, is.numeric))[11:15])
hist.data.frame(na.omit(select_if(weather, is.numeric))[16:20])
hist.data.frame(na.omit(select_if(weather, is.numeric))[21:22])
```


We can replace missing values using mean or median (there are another approaches to replace NAs, but in this homework I will use only these two options for continous data).

If column contains outliers, then mean is not the appropriate choise, and median will be more representative.

For each numerical column, excluding Day, I check whether column contains >= 1 outlier or not. If column contains >= 1 outlier, NA values are replaced by median, in other case NA values are replaced by mean.

Outliers checking is based on quantile definition and finding observations which are not present in 2*IQR interval for specific column.

```{r}
weather <- select(weather, period, Day, Events) %>% cbind(select_if(weather, is.numeric) %>% select(-Day) %>% transmute_all(replaceNA))
tbl_df(weather)
lapply(weather, function(x) sum(is.na(x)))
```

Events is a categorical data, and the mode can be used as alternative for NA values

```{r}
weather$Events <- replace(weather$Events, is.na(weather$Events), getMode(weather$Events))
```